---
title: "Homework 1"
author: "Kaushika Potluri"
desription: "Something to describe what I did"
date: "08/02/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - challenge1
  - my name
  - dataset
  - ggplot2
---

```{r}
#| label: setup
#| warning: false

knitr::opts_chunk$set(echo = TRUE)
```

### Loading in packages:
```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(readxl)

```

### Reading in Data:
```{r}
df <- read_excel('_data/LungCapData.xls')
df
```

## 1(a) Distribution of LungCap:

```{r}
hist(df$LungCap)
```
The distribution appears to be very similar to a normal distribution, according to the histogram.

## 1(b)
The boxplots below show the probability distributions grouped by Gender.
```{r}
boxplot(LungCap~Gender, data=df)
```
Looks like males have a slightly higher lung capacity than females.

## 1 (c)

```{r}
df %>%
  group_by(Smoke) %>%
  summarize(Mean = mean(LungCap))
```

Surprisingly, the mean lung capacity is higher for smokers than it is for non-smokers.

## 1 (d)
```{r}

# convert Age to categorical variable.
df <- mutate(df, AgeGroup = case_when(Age <= 13 ~ "13 and lower", Age == 14 | Age == 15 ~ "14-15", Age == 16 | Age == 17 ~ "16-17", Age >= 18 ~ "18 and above"))
arrange(df, Age)

# construct histogram.
ggplot(df, aes(x = LungCap)) +
  geom_histogram() +
  facet_grid(AgeGroup~Smoke)
```

Majority seem to be non-smokers, and looks like non-smokers seem to have higher lung capacity.

## 1 (e)

```{r}
class(df$AgeGroup)
```

```{r}
df$AgeGroup <- as.factor(df$AgeGroup) #converting to factor

# construct table.
df %>% select(Smoke, LungCap, AgeGroup) %>% group_by(AgeGroup, Smoke) %>% summarise(mean(LungCap))
```
The mean lung capacity for smokers aged 13 and under is greater than that of non-smokers in the same age group which is different from expectation. Non-smokers have higher mean lung capacity for ages 14-15, 16-17 and 18 and above.
Either there may be an error or extreme outlier in the data for smokers aged 13 and under.

## 1 (f)

```{r}
cor(df$LungCap,df$Age)
```

```{r}
cov(df$LungCap,df$Age)
```

Lung capacity and age have a high positive correlation of 0.82, meaning that as age increases, lung capacity also does. The covariance is a little more challenging to interpret; the positive number indicates a positive association between lung capacity and age, but because covariance varies from negative infinity to infinity, it is difficult to judge the strength of the relationship. In most situations, I would choose to employ correlation.

## 2
```{r}
df1 <- c(0:4)
Inmate_count <- c(128, 434, 160, 64, 24)
IP<- data_frame(df1, Inmate_count)
```

## 2(a)
```{r}
IP <- mutate(IP, Probability = Inmate_count/sum(Inmate_count))
IP
```
```{r}
IP %>%
  filter(df1 == 2) %>%
  select(Probability)
```

The probability is about 19.75%.

## (b)
```{r}
df2 <- IP %>%
  filter(df1 < 2)
sum(df2$Probability)
```

The probability that a randomly selected inmate has fewer than 2 prior convictions is 0.6938272

## 2(c)

```{r}
df3 <- IP %>%
  filter(df1 <= 2)
sum(df3$Probability)
```

The probability that a randomly selected inmate has 2 or fewer prior convictions is 0.891358.

## 2(d)
```{r}
df4 <- IP %>%
  filter(df1 > 2)
sum(df4$Probability)
```
The probability that a randomly selected inmate has more than 2 prior convictions is 0.108642.

## 2(e)
```{r}
IP <- mutate(IP, X = df1*Probability)
expectedvalue<- sum(IP$X)
expectedvalue


```

The expected value for the number of prior convictions is 1.2864198. 
We can round this to 1.

## 2(f)
```{r}
var1 <-sum(((IP$df1-expectedvalue)^2)*IP$Probability)
var1
```

```{r}
sqrt(var1)
```

The variance and the standard deviation for prior convictions are 0.8562353 and 0.9253298 respectively.