---
title: "Final Project"
author: "Ethan Campbell"
date: "10/10/2022"
format: 
  html:
    df-print: paged
    toc: true
    code-copy: true
    code-tools: true
    css: "styles.css"
categories:
  - finalpart1
---

# Introduction

Extensive research has been done on climate change and economic changes respectively but there is not a significant amount of research about their relation towards one another. There are research papers that touch on this but in different aspects and focus more on other factors like political aspects. I would like to look a little broader and look at the difference between each climate zone and their economic differences. This can be taken with a grain of salt as there are many factors that could effect the economic situation being left out. The data was pulled from NASA's POWER data access viewer; here I pulled the data by region since pulling the whole country in one go was unavailable. Thus, I will conduct research on each region respectively and then compare the results.

::: callout-tip
## Research Questions

A. Is there a relation between climate zone and economic growth?

B. Do Southern climates have the largest economic growth?

:::

# Hypothesis

My hypothesis is that there is an impact since certain climates effect the type of labor, conditions, longevity of materials, abundance of materials, and it alters the level of difficult of human growth. This study will focus more on generating climate regions and comparing the economic status within each region and try to identify trends that are occurring. I believe the regions with the biggest economic growth would be the south with their flatter lands or the west coast since its climate is good for technological advances. 

The hypothesis will be tested as follows:

::: callout-tip
## H~0A~


Climate region differences [will not]{.underline} be statistically significant predictors of economical change.
:::

::: callout-tip
## H~1A~

Climate region differences [will]{.underline} be statistically significant predictors of economical change.
:::

::: callout-tip
## H~0A~


Southern climates [will not]{.underline} be different compared to other climate regions.
:::

::: callout-tip
## H~1A~

Southern climates [will]{.underline} be different compared to other climate regions.
:::

# Library

```{r}
library(tidyverse)
library(dplyr)
library(kableExtra)
```

# Descriptive statistics

## Reading in the data

```{r}
## problem 1 how to merge files that are this large together? do we reduce the file sizes or is there another work around
# May just need to keep regions separated 
# Weather data
# Reading in all the weather data 
Amherst <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/amherst.csv", skip = 14)
Florida <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/flordia.csv", skip = 14)
Illinois <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/illinois.csv", skip = 14)
Middle <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/middle.csv", skip = 14)
Newmexico <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/Newmexico.csv", skip = 14)
North <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/North.csv", skip = 14)
South <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/south.csv", skip = 14)
SouthCali <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/SouthCali.csv", skip = 14)
Texas <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/Texas.csv", skip = 14)
Washington <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/washington.csv", skip = 14)
WestV <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/WestV.csv", skip = 14)

```

### Amherst

Had trouble with pivot_wider since it would split the values up by each name but then would fill in the values with NA for the other sections. This added a ton of NA values that one looked bad and were hard to deal with. I had to go a more manual way and do it for each part of PARAMETER to get the exact number of rows I needed. This stopped the NA values and got them all lined up so it reduced the size of the document from 500k+ rows to 89290 rows. This is huge in terms of running the data and working with it. Finally, I just merged the data together and then I was able to rename all the columns and start regression analysis.

```{r}

# Bringing all the month columns into one column
Amherst <- Amherst %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)

## getting the parameter section broken up into different sections since the normal pivot_longer did not work
Tidy_amherst <- Amherst %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_amherst <- Tidy_amherst %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_amherst <- Tidy_amherst %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)


t2m <- Amherst %>%
  select(PARAMETER, Month_Average, YEAR) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_amherst$T2M <- t2m$T2M

rh2m <- Amherst %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_amherst$RH2M <- rh2m$RH2M

wh10m <- Amherst %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_amherst$WS10M <- wh10m$WS10M


wh50m <- Amherst %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_amherst$WS50M <- wh50m$WS50M

PRECTOTCORR <- Amherst %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_amherst$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR


# renaming all the variables to easier to digest names
Tidy_amherst <- Tidy_amherst %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

## Getting them in clean looking order
Tidy_amherst <- Tidy_amherst %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

## Summary and starting regression information 
summary(Tidy_amherst)

view_amherst <- Tidy_amherst %>%
  slice(1:10)

kable(view_amherst, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)

```

### Florida

```{r}

Florida <- Florida %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)


Tidy_Florida <- Florida %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_Florida <- Tidy_Florida %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_Florida <- Tidy_Florida %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- Florida %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_Florida$T2M <- t2m$T2M

rh2m <- Florida %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_Florida$RH2M <- rh2m$RH2M

wh10m <- Florida %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_Florida$WS10M <- wh10m$WS10M


wh50m <- Florida %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_Florida$WS50M <- wh50m$WS50M

PRECTOTCORR <- Florida %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_Florida$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_Florida <- Tidy_Florida %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_Florida <- Tidy_Florida %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_Florida)

Tidy_Florida <- Tidy_Florida %>%
  slice(1:10)

kable(Tidy_Florida, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Florida Data") %>%
  kable_styling(font_size = 16)

```

### Illinois

```{r}


Illinois <- Illinois %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)


Tidy_Illinois <- Illinois %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_Illinois <- Tidy_Illinois %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_Illinois <- Tidy_Illinois %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- Illinois %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_Illinois$T2M <- t2m$T2M

rh2m <- Illinois %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_Illinois$RH2M <- rh2m$RH2M

wh10m <- Illinois %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_Illinois$WS10M <- wh10m$WS10M


wh50m <- Illinois %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_Illinois$WS50M <- wh50m$WS50M

PRECTOTCORR <- Illinois %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_Illinois$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_Illinois <- Tidy_Illinois %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_Illinois <- Tidy_Illinois %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_Illinois)

Tidy_Illinois <- Tidy_Illinois %>%
  slice(1:10)

kable(Tidy_Illinois, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### Middle 


```{r}
Middle <- Middle %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)


Tidy_Middle <- Middle %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_Middle <- Tidy_Middle %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_Middle <- Tidy_Middle %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- Middle %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_Middle$T2M <- t2m$T2M

rh2m <- Middle %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_Middle$RH2M <- rh2m$RH2M

wh10m <- Middle %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_Middle$WS10M <- wh10m$WS10M


wh50m <- Middle %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_Middle$WS50M <- wh50m$WS50M

PRECTOTCORR <- Middle %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_Middle$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_Middle <- Tidy_Middle %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_Middle <- Tidy_Middle %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_Middle)

Tidy_Middle <- Tidy_Middle %>%
  slice(1:10)

kable(Tidy_Middle, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### New Mexico

```{r}
Newmexico <- Newmexico %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)


Tidy_Newmexico <- Newmexico %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_Newmexico <- Tidy_Newmexico %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_Newmexico <- Tidy_Newmexico %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- Newmexico %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_Newmexico$T2M <- t2m$T2M

rh2m <- Newmexico %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_Newmexico$RH2M <- rh2m$RH2M

wh10m <- Newmexico %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_Newmexico$WS10M <- wh10m$WS10M


wh50m <- Newmexico %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_Newmexico$WS50M <- wh50m$WS50M

PRECTOTCORR <- Newmexico %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_Newmexico$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_Newmexico <- Tidy_Newmexico %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_Newmexico <- Tidy_Newmexico %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_Newmexico)

Tidy_Newmexico <- Tidy_Newmexico %>%
  slice(1:10)

kable(Tidy_Newmexico, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### North
```{r}
North <- North %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)
Tidy_North <- North %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_North <- Tidy_North %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_North <- Tidy_North %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)


t2m <- North %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_North$T2M <- t2m$T2M

rh2m <- North %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_North$RH2M <- rh2m$RH2M

wh10m <- North %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_North$WS10M <- wh10m$WS10M


wh50m <- North %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_North$WS50M <- wh50m$WS50M

PRECTOTCORR <- North %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_North$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_North <- Tidy_North %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_North <- Tidy_North %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_North)

Tidy_North <- Tidy_North %>%
  slice(1:10)

kable(Tidy_North, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### South

```{r}

South <- South %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)
Tidy_South <- South %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_South <- Tidy_South %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_South <- Tidy_South %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- South %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_South$T2M <- t2m$T2M

rh2m <- South %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_South$RH2M <- rh2m$RH2M

wh10m <- South %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_South$WS10M <- wh10m$WS10M


wh50m <- South %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_South$WS50M <- wh50m$WS50M

PRECTOTCORR <- South %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_South$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_South <- Tidy_South %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_South <- Tidy_South %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_South)

Tidy_South <- Tidy_South %>%
  slice(1:10)

kable(Tidy_South, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### South California
```{r}

SouthCali <- SouthCali %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)
Tidy_SouthCali <- SouthCali %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_SouthCali <- Tidy_SouthCali %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_SouthCali <- Tidy_SouthCali %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- SouthCali %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_SouthCali$T2M <- t2m$T2M

rh2m <- SouthCali %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_SouthCali$RH2M <- rh2m$RH2M

wh10m <- SouthCali %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_SouthCali$WS10M <- wh10m$WS10M


wh50m <- SouthCali %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_SouthCali$WS50M <- wh50m$WS50M

PRECTOTCORR <- SouthCali %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_SouthCali$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_SouthCali <- Tidy_SouthCali %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_SouthCali <- Tidy_SouthCali %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_SouthCali)

Tidy_SouthCali <- Tidy_SouthCali %>%
  slice(1:10)

kable(Tidy_SouthCali, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### Texas

```{r}

Texas <- Texas %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)
Tidy_Texas <- Texas %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_Texas <- Tidy_Texas %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_Texas <- Tidy_Texas %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- Texas %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_Texas$T2M <- t2m$T2M

rh2m <- Texas %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_Texas$RH2M <- rh2m$RH2M

wh10m <- Texas %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_Texas$WS10M <- wh10m$WS10M


wh50m <- Texas %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_Texas$WS50M <- wh50m$WS50M

PRECTOTCORR <- Texas %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_Texas$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_Texas <- Tidy_Texas %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_Texas <- Tidy_Texas %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_Texas)

Tidy_Texas <- Tidy_Texas %>%
  slice(1:10)

kable(Tidy_Texas, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### Washington

```{r}

Washington <- Washington %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)
Tidy_Washington <- Washington %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_Washington <- Tidy_Washington %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_Washington <- Tidy_Washington %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- Washington %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_Washington$T2M <- t2m$T2M

rh2m <- Washington %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_Washington$RH2M <- rh2m$RH2M

wh10m <- Washington %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_Washington$WS10M <- wh10m$WS10M


wh50m <- Washington %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_Washington$WS50M <- wh50m$WS50M

PRECTOTCORR <- Washington %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_Washington$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_Washington <- Tidy_Washington %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_Washington <- Tidy_Washington %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_Washington)

Tidy_Washington <- Tidy_Washington %>%
  slice(1:10)

kable(Tidy_Washington, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### West Virgina

```{r}

WestV <- WestV %>%
pivot_longer(
  cols = c(NOV, JAN, FEB, MAR, APR, MAY, JUN, JUL, AUG, SEP, OCT, DEC),
  names_to = "MONTH",
  values_to = "Month_Average",
)
Tidy_WestV <- WestV %>%
  select(PARAMETER, Month_Average, YEAR, LAT, LON, MONTH, ANN) %>%
  filter(PARAMETER == 'PS')

Tidy_WestV <- Tidy_WestV %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

Tidy_WestV <- Tidy_WestV %>%
  select(PS, YEAR, MONTH, LAT, LON, ANN)

t2m <- WestV %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'T2M')

t2m <- t2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

t2m <- t2m %>%
  select(T2M, YEAR)

Tidy_WestV$T2M <- t2m$T2M

rh2m <- WestV %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'RH2M')

rh2m <- rh2m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

rh2m <- rh2m %>%
  select(RH2M, YEAR)

Tidy_WestV$RH2M <- rh2m$RH2M

wh10m <- WestV %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS10M')

wh10m <- wh10m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh10m <- wh10m %>%
  select(WS10M, YEAR)

Tidy_WestV$WS10M <- wh10m$WS10M


wh50m <- WestV %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'WS50M')

wh50m <- wh50m %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

wh50m <- wh50m %>%
  select(WS50M, YEAR)

Tidy_WestV$WS50M <- wh50m$WS50M

PRECTOTCORR <- WestV %>%
  select(PARAMETER, Month_Average, YEAR, LAT) %>%
  filter(PARAMETER == 'PRECTOTCORR')

PRECTOTCORR <- PRECTOTCORR %>%
  group_by(PARAMETER) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = PARAMETER, values_from = Month_Average) %>%
  select(-row)

PRECTOTCORR <- PRECTOTCORR %>%
  select(PRECTOTCORR, YEAR)

Tidy_WestV$PRECTOTCORR <- PRECTOTCORR$PRECTOTCORR

Tidy_WestV <- Tidy_WestV %>%
  dplyr::rename(Temperature = T2M) %>%
  dplyr::rename(Humidity = RH2M) %>%
  dplyr::rename(Wind_10_meter = WS10M) %>%
  dplyr::rename(Surface_Pressure = PS) %>%
  dplyr::rename(Wind_50_meter = WS50M) %>%
  dplyr::rename(Precipitation = PRECTOTCORR) %>%
  dplyr::rename(Annual = ANN) %>%
  dplyr::rename(Latitude = LAT) %>%
  dplyr::rename(Longitude = LON) %>%
  dplyr::rename(Month = MONTH) %>%
  dplyr::rename(Year = YEAR)

Tidy_WestV <- Tidy_WestV %>%
  select(Year, Month, Latitude, Longitude, Temperature, Humidity, Precipitation, Surface_Pressure, Wind_10_meter, Wind_50_meter, Annual)

summary(Tidy_WestV)

Tidy_WestV <- Tidy_WestV %>%
  slice(1:10)

kable(Tidy_WestV, digits = 2, align = "ccccccc", col.names = c("Year", "Month", "Latitude", "Longitude", "Temperature", "Humidity", "Precipitation", "Surface Pressure", "Wind 10 Meters", "Wind 50 Meters", "Annual"), caption = "Amherst Data") %>%
  kable_styling(font_size = 16)
```

### State Economy data

The economic data is pulled from the Bureau of Economic Analysis (Analysis, n.d.). This data is the ins, outs, and the difference between the former two in income by state. The data ranges from 1990 to 2020 and covers every state in the US.

```{r}

# Reading in economic data

Economy <- read.csv("C:/Users/ethan/Documents/Github Class/603_Fall_2022_homework/Economy.csv")


# Renaming the columns to remove the X
Economy <- Economy %>%
  dplyr::rename('1990' = X1990) %>%
  dplyr::rename('1991' = X1991) %>%
  dplyr::rename('1992' = X1992) %>%
  dplyr::rename('1993' = X1993) %>%
  dplyr::rename('1994' = X1994) %>%
  dplyr::rename('1995' = X1995) %>%
  dplyr::rename('1996' = X1996) %>%
  dplyr::rename('1997' = X1997) %>%
  dplyr::rename('1998' = X1998) %>%
  dplyr::rename('1999' = X1999) %>%
  dplyr::rename('2000' = X2000) %>%
  dplyr::rename('2001' = X2001) %>%
  dplyr::rename('2002' = X2002) %>%
  dplyr::rename('2003' = X2003) %>%
  dplyr::rename('2004' = X2004) %>%
  dplyr::rename('2005' = X2005) %>%
  dplyr::rename('2006' = X2006) %>%
  dplyr::rename('2007' = X2007) %>%
  dplyr::rename('2008' = X2008) %>%
  dplyr::rename('2009' = X2009) %>%
  dplyr::rename('2010' = X2010) %>%
  dplyr::rename('2011' = X2011) %>%
  dplyr::rename('2012' = X2012) %>%
  dplyr::rename('2013' = X2013) %>%
  dplyr::rename('2014' = X2014) %>%
  dplyr::rename('2015' = X2015) %>%
  dplyr::rename('2016' = X2016) %>%
  dplyr::rename('2017' = X2017) %>%
  dplyr::rename('2018' = X2018) %>%
  dplyr::rename('2019' = X2019) %>%
  dplyr::rename('2020' = X2020) 

# Pivoting to combine all the years into one column
Economy <- Economy %>%
pivot_longer(
  cols = c('1990', '1991', '1992', '1993', '1994', '1995', '1996', '1997', '1998', '1999', '2000', '2001', '2002', '2003', '2004', '2005', '2006', '2007', '2008', '2009', '2010', '2011', '2012', '2013', '2014', '2015', '2016', '2017', '2018', '2019', '2020'),
  names_to = "Year",
  values_to = "Yearly_Fianace",
)

## Change from char to numeric
Economy$Year <- as.numeric(Economy$Year)

# Changing the finance column to be in millions
Economy <- Economy %>%
  mutate(Year_Money_Millions = Yearly_Fianace/1000)

summary(Economy)

```


# Bibliography 

 * Analysis, B. o. (n.d.). Regional Economic Accounts: Download. Retrieved from BEA: https://apps.bea.gov/regional/downloadzip.cfm


 * Research, NASA Langley. The POWER Project. 08 May 2021. 21 February 2022. https://power.larc.nasa.gov.

(“These data were obtained from the NASA Langley Research Center (LaRC) POWER Project funded through the NASA Earth Science/Applied Science Program.”)



