---
title: "Final project part 1"
author: "Mani Shanker Kamarapu"
description: "The first part of final project"
date: "10/11/2022"
format:
  html:
    df-print: paged
    css: styles.css
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart1
  - Bank Customer Churn Prediction
  - Mani Shanker Kamarapu
editor: 
  markdown: 
    wrap: 72
---

## Introduction

Churning refers to a customer who leaves one company to go to another
company. Customer churn introduces not only some loss in income but also
other negative effects on the operation of companies. Churn management
is the concept of identifying those customers who are intending to move
their custom to a competing service provider.

Risselada et al. (2010) stated that churn management is becoming part of
customer relationship management. It is important for companies to
consider it as they try to establish long-term relationships with
customers and maximize the value of their customer base.

::: callout-important
## Research Questions

A. Does churn-rate depend on the geographical factors of the customer?

B. Do non-active members are probable to churn or not?
:::

This project will be useful to better understand more about the customer
difficulties and factors and also give us a pretty good idea on the
factors effecting the customers to exit and also about the dormant state
of the customers.

## Hypothesis

Customer churn analysis has become a major concern in almost every
industry that offers products and services. The model developed will
help banks identify clients who are likely to be churners and develop
appropriate marketing actions to retain their valuable clients. And this
model also supports information about similar customer group to consider
which marketing reactions are to be provided. Thus, due to existing
customers are retained, it will provide banks with increased profits and
revenues.

Given the above, we can frame our hypotheses as follows:

::: callout-tip
## H~0A~

Geographical factors [will not]{.underline} be statistically predict the
churn-rate.
:::

::: callout-tip
## H~1A~

Geographical factors [will]{.underline} be statistically predict the
churn-rate.
:::

::: callout-tip
## H~0B~

Active members [will not]{.underline} churn.
:::

::: callout-tip
## H~1B~

Active members [will]{.underline} churn.
:::

## Loading libraries

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(ggplot2)
library(stats)

knitr::opts_chunk$set(echo = TRUE)
```

## Reading the data set

```{r}
Churn <- read_csv("_data/Churn_Modelling.csv")
Churn
```

This data set includes 10k bank customer data records with 14 attributes
including socio-demographic attributes, account level and behavioural
attributes.

*Attribute Description* 1. Row Number- Number of customers 2. Customer
ID- ID of customer 3. Surname- Customer name 4. Credit Score- Score of
credit card usage 5. Geography- Location of customer 6. Gender- Customer
gender 7. Age- Age of Customer 8. Tenure- The period of having the
account in months 9. Balance- Customer main balance 10. NumOfProducts-
No of products used by customer 11. HasCrCard- If the customer has a
credit card or not 12. IsActiveMember- Customer account is active or not
13. Estimated Salary- Estimated salary of the customer. 14. Exited-
Indicate churned or not

```{r}
str(Churn)
```

## Descriptive statistics

```{r}
summary(Churn)
```

```{r}
glimpse(Churn)
```
