---
title: "Final Project Part 2"
author: "Emily Duryea"
desription: "Final Project Part 2"
date: "11/11/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - final project
  - Emily Duryea
  - dataset
---

# Final Project Part 2

```{r}
studentsurvey <- read.csv("student_prediction.csv")
summary(studentsurvey)
library(ggplot2)
library(tidyverse)
```

## Research Question 1

Question: Does classroom engagement (i.e., taking notes, attending class, listening) result in a higher GPA in university students?

Hypothesis: Classroom engagement factors (i.e., taking notes, attending class, listening) will have a positive correlation with higher GPA.

Explanatory Variable: Classroom engagement factors (1) taking notes, 2) attending class, 3) listening)

Response Variable: University students' GPA

### Simple Linear Regression & Correlation Test

Notes & GPA

```{r}
nfit <- lm(NOTES ~ CUML_GPA, data = studentsurvey)
summary(nfit)

cor.test(studentsurvey$NOTES, studentsurvey$CUML_GPA)
```

Attendance & GPA

```{r}
afit <- lm(ATTEND ~ CUML_GPA, data = studentsurvey)
summary(afit)

cor.test(studentsurvey$ATTEND, studentsurvey$CUML_GPA)
```

Listening & GPA

```{r}
lfit <- lm(LISTENS ~ CUML_GPA, data = studentsurvey)
summary(lfit)

cor.test(studentsurvey$LISTENS, studentsurvey$CUML_GPA)
```

### Multiple Regression with all 3 Variables

```{r}
summary(lm(CUML_GPA ~ NOTES + ATTEND + LISTENS, data = studentsurvey))
```

### Visualizing the Data

Impact of Taking Notes on University Student GPA

```{r}
ggplot(data = studentsurvey, aes(x = NOTES, y = CUML_GPA)) +
  geom_point() +
  geom_smooth(method = lm)
```

Impact of Class Attendance on GPA

```{r}
ggplot(data = studentsurvey, aes(x = ATTEND, y = CUML_GPA)) +
  geom_point() +
  geom_smooth(method = lm)
```

Impact of Listening in Class on GPA

```{r}
ggplot(data = studentsurvey, aes(x = LISTENS, y = CUML_GPA)) +
  geom_point() +
  geom_smooth(method = lm)
```

### Conclusions

To analyze the influence of classroom engagement on student GPA, I chose to run a simple linear regression and a correlation test. I did also conduct a multiple regression analysis, but I preferred to separate the three variables within my definition of "classroom engagement" so I could analyze them individually.

Taking notes did not appear to have a significant impact on cumulative GPA. The p-value (0.08499) was greater than 0.05, indicating the result was not statistically significant. Additionally, the correlation coefficient was positive, but only slightly (0.1435413). The adjusted r squared also indicated a low correlation (0.01376).

Class attendance was found to be a statistically significant, as the p-value was less than 0.05 (0.0319). Interestingly, classroom attendance actually had a negative correlation with GPA, indicating that students who attended class less frequently obtained higher GPAs. This correlation is also slight, as indicated by the correlation coefficient (-0.1783047) and the adjusted r squared (0.02502).

Students' reported listening during class was not statistically significant on GPA, with a p-value higher than 0.05 (0.5079). The correlation was also extremely slight, with a positive correlation coefficient of 0.05542742 and an adjusted r squared value of -0.003899.

Thus, my hypothesis that classroom engagement would have a positive influence on GPA would be rejected.

## Research Question 2

Question: Does reported studying (i.e., weekly study hours) result in a higher GPA in university students?

Hypothesis: More hours studied will have a positive impact on student cumulative GPA.

Explanatory Variable: Hours reported studying a week

Response Variable: Cumulative GPA

### Simple Linear Regression & Correlation Test

```{r}
shfit <- lm(STUDY_HRS ~ CUML_GPA, data = studentsurvey)
summary(shfit)

cor.test(studentsurvey$STUDY_HRS, studentsurvey$CUML_GPA)
```

### Visualizing the Data

```{r}
ggplot(data = studentsurvey, aes(x = STUDY_HRS, y = CUML_GPA)) +
  geom_point() +
  geom_smooth(method = lm)
```

### Conclusions

Like with my previous research question, I chose to run a simple linear regression and a correlation test to analyze the data. The results indicated that hours spent studying had very little impact on cumulative GPA. The p-value was greater than 0.05 (0.9225), and both the correlation coefficient, although positive, and the adjusted r r squared values were extremely small (0.008144991 and -0.006926). Thus, my hypothesis would be refuted.

## Research Question 3

Question: Does collaboration between students (i.e., studying together, positive class discussions) result in a higher GPA in university students?

Hypothesis: Collaboration between students would have a positive impact on cumulative GPA.

Explanatory Variable: Collaboration (Studying with peers, perceiving classroom discussions as positive)

Response Variable: Cumulative GPA

### Simple Linear Regression & Correlation Test

Studying with peers & GPA

```{r}
studentsurvey$PREP_STUDY <- ifelse(studentsurvey$PREP_STUDY==2, 2, 1)

spfit <- lm(PREP_STUDY ~ CUML_GPA, data = studentsurvey)
summary(spfit)

cor.test(studentsurvey$PREP_STUDY, studentsurvey$CUML_GPA)
```

Positive discussions & GPA

```{r}
studentsurvey$PREP_STUDY <- ifelse(studentsurvey$LIKES_DISCUSS==1, 1, 2)

ldfit <- lm(LIKES_DISCUSS ~ CUML_GPA, data = studentsurvey)
summary(ldfit)

cor.test(studentsurvey$LIKES_DISCUSS, studentsurvey$CUML_GPA)
```

### Multiple Regression Analysis

```{r}
summary(lm(CUML_GPA ~ PREP_STUDY + LIKES_DISCUSS, data = studentsurvey))
```

### Visualizing the Data

Studying with peers & GPA

```{r}
ggplot(data = studentsurvey, aes(x = PREP_STUDY, y = CUML_GPA)) +
  geom_point() +
  geom_smooth(method = lm)
```

Positive discussions & GPA

```{r}
ggplot(data = studentsurvey, aes(x = LIKES_DISCUSS, y = CUML_GPA)) +
  geom_point() +
  geom_smooth(method = lm)
```

### Conclusions

Students who study with their peers are more likely to have higher GPAs, according to the simple linear regression and correlation test. The p-value was less than 0.05 (0.01535). However, the correlation was not extremely high (0.2009882) and neither was the adjusted r-squared value (0.03369). That being said, the results were statistically significant.

Additionally, students who found class discussions to be helpful (always or some of the time, compared to those who did not find class discussions to be a positive experience) to their education and learning were significantly more likely to have higher GPAs. The p-value was less than 0.01 (0.007804). Again the correlation was not extreme (0.2201251) as well as the adjusted r-squared (0.0418).

The multiple regression analysis also found the combined two variables to be statistically significant (0.01666). Thus, it could be concluded that collaboration has a positive impact on GPA, supporting my hypothesis.
