---
title: "Final project part 2"
author: "Mani Shanker Kamarapu"
description: "The second part of final project"
date: "11/11/2022"
format:
  html:
    df-print: paged
    css: styles.css
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart2
  - Bank Customer Churn Prediction
  - Mani Shanker Kamarapu
---

## Introduction

Churning refers to a customer who leaves one company to go to another
company. Customer churn introduces not only some loss in income but also
other negative effects on the operation of companies. Churn management
is the concept of identifying those customers who are intending to move
their custom to a competing service provider.

Risselada et al. (2010) stated that churn management is becoming part of
customer relationship management. It is important for companies to
consider it as they try to establish long-term relationships with
customers and maximize the value of their customer base.

::: callout-important
## Research Questions

A. Does churn-rate depend on the geographical factors(Customer's location) of the customer?

B. Do non-active members are probable to churn or not?
:::

This project will be useful to better understand more about the customer
difficulties and factors and also give us a pretty good idea on the
factors effecting the customers to exit and also about the dormant state
of the customers.

## Hypothesis

Customer churn analysis has become a major concern in almost every industry that offers products and services. The model developed will help banks identify clients who are likely to be churners and develop appropriate marketing actions to retain their valuable clients. And this model also supports information about similar customer group to consider which marketing reactions are to be provided. Thus, due to existing customers are retained, it will provide banks with increased profits and revenues. By the end of this article, let’s attempt to solve some of the key business challenges pertaining to customer attrition like say, (1) what is the likelihood of an active customer leaving an organization? (2) what are key indicators of a customer churn? (3) what retention strategies can be implemented based on the results to diminish prospective customer churn?

Given the above, we can frame our hypotheses as follows:

::: callout-tip
## H~0A~

Customer's location [will not]{.underline} be statistically predict the
churn-rate.
:::

::: callout-tip
## H~1A~

Customer's location [will]{.underline} be statistically predict the
churn-rate.
:::

I believe that the customer's location have an effect on customer's churn rate as based on location there is statistical difference in customer's salary and balance.

::: callout-tip
## H~0B~

Active members [will not]{.underline} churn.
:::

::: callout-tip
## H~1B~

Active members [will]{.underline} churn.
:::

I think that inactive members are more likely to exit rather than active members as there is a high chance of them churning out as they are are inactive for a longtime.

## Loading libraries

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(ggplot2)
library(stats)
library(reshape2)
library(skimr)
library(randomForest)
library(caret)
library(interactions)
library(lmtest)
library(sandwich)
library(plotly)

knitr::opts_chunk$set(echo = TRUE)
```

## Reading the data set

```{r}
Churn <- read_csv("_data/Churn_Modelling.csv")
Churn
```

This data set is originated from a U.S. bank and is downloaded from kaggle. This data set includes 10k bank customer data records with 14 attributes including socio-demographic attributes, account level and behavioral attributes.

## Attribute Description
1. Row Number- Number of customers 
2. Customer ID- ID of customer 
3.Surname- Customer name 
4. Credit Score- Score of credit card usage 
5. Geography- Location of customer 
6. Gender- Customer gender 
7. Age- Age of Customer 
8. Tenure- The period of having the account in months 
9. Balance- Customer main balance 
10. NumOfProducts- No of products used by customer(No of accounts the customer have)
11. HasCrCard- If the customer has a credit card or not 
12. IsActiveMember- Customer account is active or not(if he haven't used his savings or current account for any transactions for over 1 year, then he is treated as inactive.) 
13. Estimated Salary- Estimated salary of the customer. 
14. Exited- Indicate churned or not, i.e, if the customer left the bank or not.

The response variable is `Exited` variable and the main explanatory variables are `Geography` and `IsActiveMember`. And the other explanatory variables are Credit Score, Gender, Age and Balance. 

```{r}
str(Churn)
```

## Descriptive statistics

```{r}
summary(Churn)
```

```{r}
glimpse(Churn)
```

## Tidying the data

```{r}
Churn <- Churn %>%
  select(-c(RowNumber, CustomerId, Surname))
Churn
```

Dimensions of the data set

```{r}
dim(Churn)
```

The data set has 10000 rows and 11 columns now after removing the first 3 columns which are not necessary for analysis and will not effect the model.

Checking for Null values

```{r}
apply(is.na(Churn), MARGIN = 2, FUN = sum)
```

## Detecting the outliers

```{r}
Churn %>%
  ggplot(aes(CreditScore)) +
  geom_boxplot() +
  coord_flip()
```

From above box plot, Credit score variable has few outliers, but they cannot potentially affect the data set.

```{r}
Churn %>%
  ggplot(aes(Age)) +
  geom_boxplot() +
  coord_flip()
```

From above box plot, age variable has outliers (age group above 60 constitutes outliers), however there are few outliers. But they cannot potentially affect the data set.

```{r}
Churn %>%
  ggplot(aes(Tenure)) +
  geom_boxplot() +
  coord_flip()
```

From above box plot, Tenure variable has no outliers.

```{r}
Churn %>%
  ggplot(aes(Balance)) +
  geom_boxplot() +
  coord_flip()
```

From above box plot, Balance variable has no outliers.

```{r}
Churn %>%
  ggplot(aes(NumOfProducts)) +
  geom_boxplot() +
  coord_flip()
```

From above box plot, NumofProducts variable has no outliers.

```{r}
Churn %>%
  ggplot(aes(EstimatedSalary)) +
  geom_boxplot() +
  coord_flip()
```

From above box plot, EstimatedSalary variable has no outliers.

## Visualing and interpreting the variables

```{r}
Churn %>%
  ggplot(aes(CreditScore)) +
  geom_density(color="Green", alpha=0.8) +
  ggtitle("Credit score of customers") + 
  theme_classic()
```

The credit score is looking to be normal with median in range of 650-700.

```{r}
Churn %>% 
  group_by(Geography) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc)) %>%
  ggplot(aes(x = "", y = perc, fill = Geography)) +
  ggtitle("Location of customers") +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = c(1, "white", "white"),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
  guides(fill = guide_legend(title = "Geography")) +
  scale_fill_viridis_d() +
  coord_polar(theta = "y") + 
  theme_void()
  scale_fill_brewer(palette="Set1")
```

The Geography variable consists of 3 values, i.e, France(50%), Germany(25%) and Spain(25%).

```{r}
p <- Churn %>%
  ggplot() +
  geom_bar(aes(Gender)) +
  ggtitle("Gender of customers") +
  theme_classic()
ggplotly(p)
```

The Gender variable consists of Male and Female values and male count(5457) is more than female count(4543).

```{r}
p <- Churn %>%
  ggplot() +
  geom_bar(aes(Tenure)) +
  theme_classic() +
  ggtitle("No of customers over their tenure")
ggplotly(p)
```

The tenure of all customers is between 0-10 years and is almost equal no of customers in each year.

```{r}
p <- Churn %>%
  filter(Balance != 0) %>%
  ggplot(aes(Balance)) +
  geom_histogram(col = "white") +
  theme_classic() +
  ggtitle("Balance of customers")
ggplotly(p)
```

We have a lot of people with balance as zero but if we ignore that the other values form a normal at 120000 in median.

```{r}
Churn %>% 
  group_by(NumOfProducts) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc)) %>%
  ggplot(aes(x = "", y = perc, fill = as.factor(NumOfProducts))) +
  ggtitle("No of products owned by customers") +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = c(1, "white", "white", "white"),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
  guides(fill = guide_legend(title = "NumofProducts")) +
  scale_fill_viridis_d() +
  coord_polar(theta = "y") + 
  theme_void()
  scale_fill_brewer(palette="Set1")
```

According to above plot, the maximum no of the products owned by customers is 4 and minimum is 1. Majority of customers own either 1  or 2 products.

```{r}
p <- Churn %>%
  ggplot() +
  geom_bar(aes(HasCrCard)) +
  ggtitle("No of customers having credit card") +
  theme_classic()
ggplotly(p)
```

According to above plot, 7055 customers have credit card and 2945 customers does not have credit card.

```{r}
p <- Churn %>%
  ggplot() +
  geom_bar(aes(IsActiveMember)) +
  ggtitle("Active customers") +
  theme_classic()
ggplotly(p)
```

Form the above plot, it looks like there are as many inactive members(4849) as active members(5151).

```{r}
Churn %>%
  ggplot(aes(EstimatedSalary)) +
  geom_density(color="Blue", alpha=0.8) +
  ggtitle("Estimated salary of customers") + 
  theme_classic()
```

From above graph, the data set contains the customers of all types of income from 0-200000.

```{r}
Churn %>% 
  group_by(Exited) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc)) %>%
  ggplot(aes(x = "", y = perc, fill = as.factor(Exited))) +
  ggtitle("Churn-rate of customers") +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = c(1, "white"),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
  guides(fill = guide_legend(title = "Churn-rate")) +
  scale_fill_viridis_d() +
  coord_polar(theta = "y") + 
  theme_void()
  scale_fill_brewer(palette="Set1")
```

From the pie chart, 80% of customers are not churned and 20% have already exited.

## Relationship between the variables

```{r}
temp <- Churn %>%
  select(-c(Geography, Gender))
round(cor(temp),3) %>%
  melt() %>% 
  ggplot(aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() +
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) + 
  labs(x = NULL, y = NULL) + 
  ggtitle("Correlation plot") +
  theme(axis.text.x = element_text(angle = 90))
```

Churn has a positive correlation with age, balance and estimated salary. Generally the correlation coefficients are not so high.

## Relationship between churn-rate and categorical variables

There are 4 categorical variables in the data set as follows:

```{r}
p <- Churn %>%
  group_by(Geography, Exited) %>%
  count() %>%
  ggplot(aes(fill = Exited, y = n, x = Geography)) + 
  geom_bar(position = "dodge", stat = "identity") +
  ggtitle("Churn-rate on basis of location") +
  xlab("Geography") +
  ylab("Frequency") +
  theme_classic()
ggplotly(p)
```

Majority of the data is from persons from France. However, the proportion of churned customers is with inversely related to the population of customers alluding to the bank possibly having a problem (maybe not enough customer service resources allocated) in the areas where it has fewer clients.

```{r}
p <- Churn %>%
  group_by(Gender, Exited) %>%
  count() %>%
  ggplot(aes(fill = Exited, y = n, x = Gender)) + 
  geom_bar(position = "dodge", stat = "identity") +
  ggtitle("Churn-rate per Gender") +
  xlab("Gender") +
  ylab("Frequency") +
  theme_classic()
ggplotly(p)
```

The proportion of female customers churning is also greater than that of male customers.

```{r}
p <- Churn %>%
  group_by(HasCrCard, Exited) %>%
  count() %>%
  ggplot(aes(fill = Exited, y = n, x = HasCrCard)) + 
  geom_bar(position = "dodge", stat = "identity") +
  ggtitle("Churn-rate per customer's credit card status") +
  xlab("Credit Card status") +
  ylab("Frequency") +
  theme_classic()
ggplotly(p)
```

Majority of the customers that churned are those with credit cards. Given that majority of the customers have credit cards could prove this to be just a coincidence.

```{r}
p <- Churn %>%
  group_by(IsActiveMember, Exited) %>%
  count() %>%
  ggplot(aes(fill = Exited, y = n, x = IsActiveMember)) + 
  geom_bar(position = "dodge", stat = "identity") +
  ggtitle("Churn-rate on basis of customer's activity") +
  xlab("Active Member") +
  ylab("Frequency") +
  theme_classic()
ggplotly(p)
```

Unsurprisingly the inactive members have a greater churn. Worryingly is that the overall proportion of inactive mebers is quite high suggesting that the bank may need a program implemented to turn this group to active customers as this will definately have a positive impact on the customer churn.

## Relationship between churn-rate and continuous variables

There are 6 continuous variables in the data set as follows:

```{r}
p <- Churn %>% 
  ggplot(aes(x = Exited, y = CreditScore, fill = as.factor(Exited))) + 
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("Churn-rate on basis of customer's credit score")
ggplotly(p)
```

There is no significant difference in the credit score distribution between retained and churned customers.

```{r}
p <- Churn %>% 
  ggplot(aes(x = Exited, y = Age, fill = as.factor(Exited))) + 
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("Churn-rate on basis of customer's age")
ggplotly(p)
```

The older customers are churning at more rate than the younger ones alluding to a difference in service preference in the age categories. The bank may need to review their target market or review the strategy for retention between the different age groups.

```{r}
p <- Churn %>% 
  ggplot(aes(x = Exited, y = Tenure, fill = as.factor(Exited))) + 
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("Churn-rate on basis of customer's tenure")
ggplotly(p)
```

With regard to the tenure, the clients on either extreme end (spent little time with the bank or a lot of time with the bank) are more likely to churn compared to those that are of average tenure.

```{r}
p <- Churn %>% 
  ggplot(aes(x = Exited, y = Balance, fill = as.factor(Exited))) + 
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("Churn-rate on basis of customer's balance")
ggplotly(p)
```

The bank is losing customers with significant bank balances which is likely to hit their available capital for lending.

```{r}
p <- Churn %>% 
  ggplot(aes(x = Exited, y = NumOfProducts, fill = as.factor(Exited))) + 
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("Churn-rate on basis of no of accounts customer's own")
ggplotly(p)
```

The no of products not has a significant effect on the likelihood to churn.

```{r}
p <- Churn %>% 
  ggplot(aes(x = Exited, y = EstimatedSalary, fill = as.factor(Exited))) + 
  geom_boxplot(alpha=0.3) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Dark2") +
  ggtitle("Churn-rate on basis of customer's salary")
ggplotly(p)
```

The Estimated salary not has a significant effect on the likelihood to churn.

## Regression models

```{r}
model1 <- lm(Exited ~ as.factor(IsActiveMember), data = Churn)
summary(model1)
```

```{r}
model2 <- lm(Exited ~ ., data = Churn)
summary(model2)
```

```{r}
model3 <- lm(Exited ~ Geography, data = Churn)
summary(model3)
```

Summarizing the above three models, the model1 describes the regression between Exited and Active member and secone model predicts the churn rate based on all the variables and third model is analysis between exited and geography. And model1 and model3 seems to be significant proving our hypothesis.

## Backward Elimination

```{r}
model4 <- lm(Exited ~ . -Geography -HasCrCard -NumOfProducts -EstimatedSalary -Tenure, data = Churn)
summary(model4)
```

In the model4, I have used backward elimination process by removing the highest p values to get a significant model and it is significant after removing above 5 variables removed in call.

## Model Evaluation

```{r}
par(mfrow = c(2,3)); plot(model1, which = 1:6)
par(mfrow = c(2,3)); plot(model2, which = 1:6)
par(mfrow = c(2,3)); plot(model3, which = 1:6)
par(mfrow = c(2,3)); plot(model4, which = 1:6)
```

According to the diagnostic plots, none of the models seem to fit super well. There are violations of assumptions in models. In some the residuals seem to have a trend (higher fitted values have lower residuals). Same with the Q-Q, plot, lower theoretical quantiles gave significantly lower standardized residuals. The scale location graph has a negative trend, suggesting variance may not be constant. Cooks dist to leverage 2898 has a high cooks distance and leverage and likely has a large influence on the model. The other models display similar issues.

## Further Study

Moving into part three of the project, I may look into other control variables that may improve the model or other transformations to improve R squared. I would try the logistic regression and randomforest models and check them in part3.

## Bibliography

Chicco, D. & Jurman, G., 2020. The advantages of the Matthews correlation coefficient (MCC)
over F1 score and accuracy in binary classification evaluation. BMC genomics, 21(1), pp. 6-13.

Colgate, M., Stewart, K. & Kinsella, R., 1996. Customer Defection: A study of the student market
in Ireland. International Journal of Bank Marketing, 14(3), pp. 23-29.

De Caigny, A., Coussement, K. & De Bock, K. W., 2018. A new hybrid classification algorithm
for customer churn prediction based on logistic regression and decision trees. European Journal
of Operational Research, 269(2), pp. 760-772.

Delgado, R. & Tibau, X. 2019. Why Cohen's Kappa should be avoided as performance measure
in classification, PLOS ONE, 14(9), pp. e0222916.

Ganesh, J., Arnold, M. J. & Reynolds, K. E., 2000. Understanding the Customer Base of Service
Providers: An Examination of the Differences between Switchers and Stayers. Journal of
Marketing, 64(3), pp. 65-87.

Gorodkin, J., 2004. Comparing two K-category assignments by a K-category correlation
coefficient. Computational Biology and Chemistry, 28(5), pp. 367-374.

Hair, J. F., Black, J. W. C., Babin, B. J. & Anderson, R. E., 2014. Multivariate Data Analysis. 7th ed. Harlow: Pearson international edn.

Hastie, T., Tibshirani, R. & Friedman, J., 2009. The Elements of Statistical Learning: data mining, inference, and prediction. 2nd ed. New York, NY: Springer New York.

Hosmer, D. W., Lemeshow, S. & Sturdivant, R. X., 2013. Applied logistic regression, 3rd ed. New
Jersey, NJ: Wiley.

James, G., Witten, D., Hastie, T. & Tibshirani, R., 2013. An Introduction to Statistical Learning: with Applications in R. New York, NY: Springer New York.

McHugh, M. L., 2012. Interrater reliability: the Kappa Statistic. Biochemia Medica, 22(3), pp.
276-282.

The Economist, 2019. A Whole New World: How technology is driving the evolution of intelligent
banking, London: The Economist Intelligence Unit (EIU).

Verbeke, W. et al., 2012. New insights into churn prediction in the Telecommunication Sector: A
profit driven data mining approach. European Journal of Operational Research, 218(1), pp. 211-
229.