---
title: "Final_Project_1"
author: "Mani Kanta Gogula & Rahul Gundeti"
description: "Project Proposal"
date: "10/11/2022"
format:
  html:
    df-print: paged
    css: styles.css
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Project Proposal
  - desriptive statistics 
  - probability
---

Heart Disease in the United States
Heart disease is the leading cause of death for men, women, and people of most racial and ethnic groups in the United States.1
One person dies every 34 seconds in the United States from cardiovascular disease.1
About 697,000 people in the United States died from heart disease in 2020—that’s 1 in every 5 deaths.1,2
Heart disease cost the United States about $229 billion each year from 2017 to 2018.3 This includes the cost of health care services, medicines, and lost productivity due to death.

Research Question :  examining the relationship between the maximum heart rate one can achieve during exercise and the likelihood of developing heart disease. Using multiple logistic regression, examining handle the confounding effects of age and gender.

Hypothesis Testing : Is there any statistical difference between the gender and age in terms of heart attack prediction.



#Loading Dataset

```{r}
library(readr)
library(tidyverse)
heart_cleveland_upload <- read_csv("heart_cleveland_upload.csv")
head(heart_cleveland_upload)
```
```{r}
dim(heart_cleveland_upload)
```

Data set contains 297 Columns and 14 columns

```{r}
colnames(heart_cleveland_upload)
```

here are 13 attributes

age: age in years
sex: sex (1 = male; 0 = female)
cp: chest pain type
-- Value 0: typical angina
-- Value 1: atypical angina
-- Value 2: non-anginal pain
-- Value 3: asymptomatic
trestbps: resting blood pressure (in mm Hg on admission to the hospital)
chol: serum cholestoral in mg/dl
fbs: (fasting blood sugar > 120 mg/dl) (1 = true; 0 = false)
restecg: resting electrocardiographic results
-- Value 0: normal
-- Value 1: having ST-T wave abnormality (T wave inversions and/or ST elevation or depression of > 0.05 mV)
-- Value 2: showing probable or definite left ventricular hypertrophy by Estes' criteria
thalach: maximum heart rate achieved
exang: exercise induced angina (1 = yes; 0 = no)
oldpeak = ST depression induced by exercise relative to rest
slope: the slope of the peak exercise ST segment
-- Value 0: upsloping
-- Value 1: flat
-- Value 2: downsloping
ca: number of major vessels (0-3) colored by flourosopy
thal: 0 = normal; 1 = fixed defect; 2 = reversable defect
and the label
condition: 0 = no disease, 1 = disease

# Descriptive statistics

```{r}
summary(heart_cleveland_upload)

```

```{r}
glimpse(heart_cleveland_upload)
```

