---
title: "Homework 3"
author: "Kaushika Potluri"
date: "11/07/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3installin
---


```{r, echo=T}
library(tidyverse)
library(ggplot2)
library(stats)
library(alr4)
library(smss)
knitr::opts_chunk$set(echo = TRUE)
```


##Question 1

Loading data :
```{r, echo=T}
data1 = data(UN11)
UN11
```
1.1.1

The Predicted variable here is ppgdp (GDP per capita)
The response is fertility.

1.1.2

```{r}
ggplot(data = UN11, aes(x = ppgdp, y = fertility)) +
  geom_point()
```

```{r}
ggplot(data = UN11, aes(x = log(ppgdp), y = log(fertility))) +
  geom_point()
```

Yes, linear regression seems plausible. 

##Question 2
a.

```{r}
UN11$british <- 1.33*UN11$ppgdp
# check slope.
summary(lm(fertility ~ british, UN11))
```

2b
```{r}
# correlation with US dollars.
cor(UN11$ppgdp, UN11$fertility)
```

```{r}
# correlation with British pounds.
cor(UN11$british, UN11$fertility)
```

## Question 3
```{r}
data(water)

#generating scatterplots.
pairs(water)
```
Looking at the plot, it seems that the stream run-off variable has a relationship to the ‘O’ named lakes but no real notable relationship to the ‘A’ named lakes.

## Question 4
```{r}

# load dataset.
data(Rateprof)

# create subset.
rateprof <- Rateprof %>% select(quality, helpfulness, clarity, easiness, raterInterest)

# generate scatterplots.
pairs(rateprof)
```

Quality, helpfulness and clarity have the clearest linear relationships with one another. Easiness and raterInterest do not seem to have linear relationships with the other variables.

##Question 5
a.
```{r}
# load dataset.
data(student.survey)
glimpse(student.survey)
```

```{r}
# generate plots.
boxplot(pi ~ re, student.survey)
```

```{r}
scatterplot(hi ~ tv, student.survey)
```


Religiosity and conservatism seem to have a positive relationship.
High school GPA and TV-watching seem to have a negative relationship.

5b
```{r}
# change pi to numeric variable.
student.survey$pi <- as.numeric(student.survey$pi)

# removing ordering in re and rename it.
levels(student.survey$re) <- c("N", "O", "M", "E")
student.survey$re <- factor(student.survey$re, ordered = FALSE)

# run regression models.
summary(lm(pi ~ re, student.survey))
```

```{r}
summary(lm(hi ~ tv, student.survey))
```


When compared to those who never went, those who regularly attended religious services were much more likely to be conservative, p<.001.Between those who occasionally attended religious services and those who never did, there was little variation in political orientation. 

Watching less hours of TV per week was associated with higher high-school GPAs, p < .05. That being said, as the R2 is fairly low, hours of TV watching is not a great predictor of high school GPA.
