---
title: "Homework 2 - Emily Duryea"
author: "Emily Duryea"
description: "The second homework assignment for DACSS 603"
date: "10/17/2022"
format:
  html:
  toc: true
  code-fold: true
  code-copy: true
  code-tools: true
categories:
  - hw2
  - Emily Duryea
---

# Homework 2

Uploading packages to be used for this assignment:

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
```

## Question 1

Below is the code used for setting up the degrees of freedom (df = sample size - 1). Thus, Bypass df would be 539 - 1 = 538, and, Angiography would be 847 - 1 = 846.

```{r}
Bypass_df = 538
Angio_df = 846
```

Below is the code for setting up the mean and standard deviation.

```{r}
Bypass_mean = 19
Bypass_sd = 10
Angio_mean = 18
Angio_sd = 9
```

The code for calculating t-score with 90% confidence interval is below.

```{r}
Bypass_tscore <- qt(p = 0.9, df = Bypass_df)
Angio_tscore <- qt(p = 0.9, df = Angio_df)
```

The code for calculating the standard error (sd/sqrt(sample size)) is below.

```{r}
Bypass_se <- Bypass_sd/sqrt(539)
Angio_se <- Angio_sd/sqrt(847)
```

The code for calculating the margin of error (T-score multiplied by the standard error) is below.

```{r}
Bypass_me <- Bypass_tscore*Bypass_se
Angio_me <- Angio_tscore*Angio_se
```

Below is the code for calculating the upper and lower ranges (add mean to margin of error for upper, subtract for lower).

```{r}
Bypass_low <- Bypass_mean - Bypass_me
Bypass_up <- Bypass_mean + Bypass_me
Angio_low <- Angio_mean - Angio_me
Angio_up <- Angio_mean + Angio_me
Bypass <- c(Bypass_low, Bypass_up)
Angio <- c(Angio_low, Angio_up)
Bypass
Angio
```

The 90% confidence interval for Bypass is \[18.45, 19.55\], and for Angio it is \[17.60, 18.40\]. Thus, Angio has the narrower confidence interval, which is logical to conclude because it has a larger sample size (847 \> 539), which reduces the margin of error. Additionally, the standard deviation is smaller (9\<10), which signifies less variance.

## Question 2

```{r}
# Number who believed college ed is essential for success
N <- 567

# Total sample size
S <- 1031

# Calculating point estimate
PE <- N/S

# Using the function prop.test() to find the confidence interval range & p-value
prop.test(N, S, PE)
```

The 95% confidence interval is \[0.519, 0.580\], with a p-value of 0.550.

## Question 3

```{r}
# Calculating confidence interval of 95%
CI95 <- qnorm(0.025, lower.tail = F)
# Calculating sample size needed using confidence interval equation
Student_sample <- ((170*0.25/5)*CI95)^2
Student_sample
```

Based on these calculations, the needed sample size would be 278 students for a significance level of 5%.

## Question 4

### Part A

```{r}
# Calculating the standard error (sd = 90 sample = 9)
company_se <- 90/sqrt(9)
company_se

# Calculating the t-score
company_tscore <- (410-500)/company_se
company_tscore

# Calculating the p-value (df = 9-1 = 8)
company_pvalue <- (pt(q=-3, df=8))*2
company_pvalue
```

It is possible to reject the null hypothesis, as the p-value is statistically significant (0.017), less than 0.05.

### Part B

```{r}
# Calculating the probability of a random sample with a mean of 410 or less
less_company <- pt(-3, 8)
less_company
```

The p-value for the lower tail is 0.00854.

### Part B

```{r}
# Calculating the probability of a random sample with a mean of 410 or more
more_company <- pt(-3, 8, lower.tail = F)
more_company
```

The p-value for the upper tail is 0.991.

```{r}
total_company <- less_company + more_company
total_company
```

The total of both tails is equal to 1.

## Question 5

### Part A

```{r}
# Calculating t-scores
Jones_tscore <- (519.5-500)/10
Jones_tscore
Smith_tscore <- (519.7-500)/10
Smith_tscore

# Calculating p-values
Jones_pvalue <- (pt(q=1.95, df=999, lower.tail=FALSE))*2
Jones_pvalue
Smith_pvalue <- (pt(q=1.97, df=999, lower.tail=FALSE))*2
Smith_pvalue
```

### Part B

If the significance level is 0.05, then Smith has statistically significant study findings, while Jones does not.

### Part C

Both of the studies could be statistically significant, depending on the significance level. For example, a 0.1 significance level would mean Jones also had statistically significant results. In this case, since the t-scores were so similar (0.2 off), it would not be unreasonable for both studies to reject the null hypothesis.

## Question 6

```{r}
gas_taxes <- c(51.27, 47.43, 38.89, 41.95, 28.61, 41.29, 52.19, 49.48, 35.02, 48.13, 39.28, 54.41, 41.66, 30.28, 18.49, 38.72, 33.41, 45.02)

# Getting the mean
mean_gtaxes <- mean(gas_taxes)
mean_gtaxes

# Conducting t-test
t.test(gas_taxes, mu=45.0, alternative="less")
```

Because the p-value is 0.03827 on a 95% confidence interval, on the 0.05 significance level, it is possible the null hypothesis that gas prices are equal to or greater than \$0.45.
