---
title: "Homework 4"
author: "Kaushika Potluri"
desription: "Homework 4"
date: "14/11/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
editor: 
  markdown: 
    wrap: sentence
---

## Loading Libraries:
```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(stringr)
library(alr4)
library(smss)
```
#1(a)
Prediction equation: ŷ = −10,536 + 53.8x1 + 2.84x2.
```{r}
#Plugging in the values of home and lot size into prediction equation
predicted_sellingprice <- -10536 + 53.8*1240 + 2.84*18000
predicted_sellingprice
```

$107,297 is the predicticted selling price

```{r}
#Residual = Actual-predicted
residual <- 145000-107296
residual
```

Therfore, the Residual is $37,704.

#1(b)

The prediction equation is ŷ = -10536 + 53.8x1 + 2.84x2. Using this, where x2= lot size, the house selling price is expected to increase by 53.8 dollars per each square-foot increase in home size given the lot sized is fixed. This is due to a fixed lot size would make 2.84x2 a set number in the prediction equation. Which means in the prediction equation y = -10356 + 53.8x1 + 2.84x2 x1 would have input values making it increase.

#1(c)
For fixed home size, 53.8 * 1 = 2.84x2
```{r}
x2 <- 53.8/2.84
x2
```

An increase in lot size of about 18.94 square-feet would have the same impact as an increase of 1 square-foot in home size on the predicted selling price.

#2
```{r}
data("salary")
salary
```

#2(a)
```{r}
summary(lm(salary ~ sex, data = salary))
```

Here the null hypothesis would be: mean salary for men and women are equal
The Alternative hypothesis would be: the salaries are not equal for men and women.
Here, the female coefficient is -3340, which can imply that women do make less than men not considering any other variables. However, if we consider the other variables and also there is a significance level of 0.07, hence we fail to reject the null hypothesis.
Therefore, we cannot conclude that there is a difference between mean salaries for men and women.

#2(b)
```{r}
model <- lm(salary ~ ., data = salary)
summary(model)
```

```{r}
confint(model)

```

The 95% confidence interval for the difference in salary between male and females is -697.82 and 3030.56.

#2(d)

DegreeePhD: For a faculty member that has a PhD degree their predicted salary is $1388.61 than other faculty members who don’t have a PhD degree. 

Rank: The baseline category is asst prof.For an associate professor their predicted salary is $5,292.36. For a professor their predicted salary is $11,118.76.
These salary differences are statistically significant at the 0.0001 alpha level for both Asst and Professor rank.


Sex: For a faculty member who is female their predicted salary is $1166.37 more than a male. However, his coefficient is not statistically significant at any alpha level.

Year: Every year a faculty member’s salary is expected to increase by $478.31.The coeffiticent is significant at the 0.0001 alpha level.

ysdegree: For every year after degree completion they can expect to have their slary decrese by $124.57. However this coefficient is not significant at any alpha level.

#2(d)
```{r}
salary$rank<- relevel(salary$rank, ref = 'Assoc')
summary(lm(salary ~ degree + rank + sex + year + ysdeg, data=salary))
```
The baseline category is now Assoc. According to these coefficients, faculty of rank asst are expected to make $5292.36 less than Associate professors. Faculty of rank Professor are expected to make $5826.40 more than Associate professors.

#2(e)
```{r}
summary(lm(salary ~ degree + sex + year + ysdeg, data=salary))

```

Removing the rank variable reveals a difference between male and female salaries with females making $1286.54 less than men. However, this difference is not signficant at any standard alpha levels.

#2(f)

```{r}
salary<-mutate(salary, hired= case_when(ysdeg<15 ~ "new", ysdeg>=15 ~ "old"))
summary(lm( salary ~ degree + sex + rank +  hired + year, data = salary))
```
According to this equation, faculty hired by the old dean make $2421.60 less than new faculty when we control for other factors. This is significant at the 0.05 alpha level.
Excluded ysdegree after creating the new variable to avoid multicollinearity.

#3(a)
```{r}
data(house.selling.price)
house.selling.price
```

```{r}
summary(lm(Price ~ Size + New, data= house.selling.price))
```

According to the coefficient for size, the price of a house is expected to increase by $116.132 for each square foot increase in size. The coefficient is significant at the 0.0001 alpha level, meaning there is a strong correlation between size and price when the age status (new/old) is held fixed.

According to the coefficient for new, a new house is expected to cost $57,736.283 more than an old house. This variable is significant at the 0.001 level, meaning that whether a house is old or new has a strong positive impact on price of the house.

#3(b)

Y = -40230.867 + 116.132(X1) + 57736.283 (X2) where X1 represents size and X2 represents new/old.

For a new house: Y = -40230.867 + 116.132(size) + 57736.283

For an old house: Y = -40230.867 + 116.132(size)

#3(c)
```{r}
size<- 3000
-40230.867 + (116.132 * size) + 57736.283 
```

```{r}
size<- 3000
-40230.867 + (116.132 *size)
```

#3(d)
```{r}
summary(lm( Price~ Size*New, data = house.selling.price))
```

#3(e)
The predicted selling price, based on the new regression that includes interaction between Size and Newness, would look like:

For a new house: Y = -22227.81 + 104.44 * Size - 78527.50 * 1 + 61.92 * Size * 1

Old: Y = -22227.81 + 104.44 * Size

#3(f)
```{r}
#new: 
-2227.808 + 166.354*3000 - 78527.502
```

```{r}
#not new:
-2227.808 + 104.438*3000

```

New: $418,306.70 Not new: $311,086.20

#3(g)
```{r}
#new: 
-2227.808 + 166.354*1500 - 78527.502
```

```{r}
Size <- 1500
New_Price = -22227.81 + 104.44 * Size - 78527.50 * 1 + 61.92 * Size * 1
Old_Price = -22227.81 + 104.44 * Size
New_Price
```

```{r}
Old_Price
```

As size of home goes up, the difference in predicted selling prices between old and new homes becomes larger.
Houses that are larger are much greater in price, especially when comparing new large houses to small new houses.

#3(h)

I prefer the second model with the interaction term which provides a clearer picture of how increased square footage makes a larger difference in bigger sized houses. The model with the interaction term also has a larger adjusted R squared.

```{r}
#for a 1000 sq foot home:
#New:
-2227.808 + 166.354*1000 - 78527.502
```

```{r}
#Not new:
-2227.808 + 104.438*1000
```

I wouldn't use this model for small homes: for a home that is 1000 square feet, the predicted price for a new house is greater than for an old house. 
So I dont think this model would be good at predicting tiny house prices.
