---
title: "HW 2 DACSS Eris Dodds"
format:
  html:
    theme: default
---

### Question 1
See Code

```{r}
bypass_mean<-19
bypass_sd<-10
bypass_size<-539
standard_error_bypass<-bypass_sd/sqrt(bypass_size)
confidence_level<-0.9
tail_area<-(1-confidence_level)/2
t_score<-qt(p = 1-tail_area, df = bypass_size-1)
CI_bypass<-c(bypass_mean - t_score * standard_error_bypass, bypass_mean + t_score * standard_error_bypass)
print(CI_bypass)
anio_mean<-18
anio_sd<-9
anio_size<-847
standard_error_anio<-anio_sd/sqrt(anio_size)
tail_area_anio<-(1-confidence_level)/2
t_score_anio<-qt(p = 1-tail_area_anio, df = anio_size)
CI_anio<-c(anio_mean - t_score_anio * standard_error_anio, anio_mean + t_score_anio * standard_error_anio)
print(CI_anio)
```


### Question 2
PE<-.55
Lower Limit = .52
Upper Limit = .58

```{r}
pop<-567
size<-1031
PE<-pop/size
prop.test(pop,size)
```


### Question 3

```{r}
(1.96)^2 *(42.5)^2/ (5)^2
```

### Question 4
a) T test indicates that the income for females is not equal to 500$ a week
b)

```{r}
female_salary<-c(rnorm(9, mean = 410, sd = 90))
t.test(female_salary, mu = 500)
t.test(female_salary, mu = 500, alternative = "less")
t.test(female_salary, mu = 500, alternative = "greater")
```

### Question 5
a) see code
b) Jones = .051 not significant, cannot reject null. Smith = .049 significant, reject null.
c) Being broad about the direction of the p value, in this case, would overshadow how marginally significant and insignificant the p values actually came out to in this case. 

```{r}
jones_mean<-519.5
smith_mean<-519.7
jones_se<-10
smith_se<-10
null<-500
jones_t<-(jones_mean-null)/jones_se
jones_t
smith_t<-(smith_mean-null)/smith_se
smith_t
jones_p<-pt(jones_t, df=999, lower.tail = FALSE)*2
jones_p
smith_p<-pt(smith_t, df=999, lower.tail = FALSE)*2
smith_p
```


### Question 6
The results of a t test show that the mean is less that 45, with a relatively small p value. We can have more confidence, then, that the average gas tax per gallon was less than .45 cents. 
```{r}
gas_taxes <- c(51.27, 47.43, 38.89, 41.95, 28.61, 41.29, 52.19, 49.48, 35.02, 48.13, 39.28, 54.41, 41.66, 30.28, 18.49, 38.72, 33.41, 45.02)
t.test(gas_taxes, mu = 45, alternative = "less")
```