---
title: "Homework 1"
author: "Yakub Rabiutheen"
description: "The first homework on descriptive statistics and probability"
date: "09/20/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Question 1

## a

First, let's read in the data from the Excel file:

```{r, echo=T}
library(readxl)
df <- read_excel("_data/LungCapData.xls")
```


The distribution of LungCap looks as follows:

```{r, echo=T}
hist(df$LungCap,freq = FALSE)
```

The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).




## b

Comparison of the Genders for both Men and Women using a Boxplot.

```{r}
boxplot(df$LungCap ~ df$Gender)
```

## c


Here is the capacity of Smokers vs Non-Smokers
```{r}
boxplot(df$LungCap~df$Smoke,
        ylab = "Capacity", 
        main = "Lung Capacity of Smokers Vs Non-Smokers",
        las = 1)
```





## d
Let's break it down even further, this is the Lung Capacity by Age Group


```{r}
df$Agegroups<-cut(df$Age,breaks=c(-Inf, 13, 15, 17, 20), labels=c("0-13 years", "14-15 years", "16-17 years", "18+ years"))
```



Below is the overall Lung Capacity of Age Groups without including Smokers.
```{r}
library(ggplot2)
ggplot(df, aes(x = LungCap, y = Agegroups, fill = Gender)) +
          geom_bar(stat = "identity") +
          coord_flip() +
          theme_classic()
```
#e

Here is a comparision of AgeGroup Lung Capacity in comparison with Smoker vs Non-Smoker.
```{r}
ggplot(df, aes(x = LungCap, y = Agegroups, fill = Smoke)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    theme_classic()
```
## 1f

Based on the comparison of lung capacities between Smoker and Non-Smoker the results are pretty similar.

```{r}
cov(df$LungCap, df$Age)
cor(df$LungCap, df$Age)
```



Question 2


```{r}
X <- c(0:4)
Frequency <- c(128, 434, 160, 64, 24)
df <- data.frame(X, Frequency)
df
```

As shown below, the most common Prior Convictions is 1.
```{r}
df
```

Dividing by the total among 810 we can determine the probability for each. 810 is the Sum of the Frequency which I checked manually.

```{r}
df2 <- mutate(df, Probability = Frequency/sum(Frequency))
df2
```

B.
Filter for Probability of 2 Convictions
```{r}
b2 <- df2 %>% 
  filter(X < 2)
sum(b2$Probability)
```
C.
Filter for Probability of Less than 2 Convictions
```{r}
c2 <- df2 %>% 
  filter(X <= 2)
sum(c2$Probability)
```


D.

Filter for Probability of greater than 2 convictions.

```{r}
d2 <- df2 %>% 
  filter(X > 2)
sum(d2$Probability)
```


What is the expected value of the number of prior convictions?

```{r}
e <- weighted.mean(df2$X, df2$Probability)
e
```



F.


Variance and Standard Deviation for Question.
```{r}
var(df$X)
```
```{r}
sd(df$X)
```

