---
title: "Homework 1"
author: "Ethan Campbell"
description: "The first homework on descriptive statistics and probability"
date: "09/21/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Question 1


First, let's read in the data from the Excel file:

```{r, echo=T}
library(readxl)
library(tidyverse)
library(dplyr)
df <- read_excel("_data/LungCapData.xls")
```


## 1.a 

(The distribution of LungCap looks as follows:)

The histogram suggests that the distribution is close to a normal distribution. Most of the observations are close to the mean. Very few observations are close to the margins (0 and 15).

```{r, echo=T}
head(df)
hist(df$LungCap)
```



## 1.b 

(Comparing lung cap by gender)

Here we notice that males tend to have a higher lung cap compared to females. Females average tends to sit around 8 while males seems to sit closer to 9

```{r}
boxplot(df$LungCap~df$Gender)
```

## 1.c 

(smoker vs non-smoker lung cap)

Interestingly, none smokers tend to have a lower lung capacity however, I believe this might be due to age. No this does not make sense at first glance and does betray my expectation.

```{r}
df %>%
  group_by(Smoke) %>%
  summarize_at(vars(LungCap), list(mean = mean))

```

## 1.d 

(relation between smoking and lung cap at different age groups)

The lung cap starts off higher but takes and dip then rises as the age continues to grow. I believe the trend is the higher age grows the higher the lung cap until it reaches a certain point. 

```{r}
# lung cap is 9.62
df %>%
  select(Age, LungCap) %>%
  filter(Age >= 13) %>%
  colMeans()
# lung cap is 9.04
df %>%
  select(Age, LungCap) %>%
  filter(Age >= 14 & Age <= 15) %>%
  colMeans()
# lung cap is 10.24
df %>%
  select(Age, LungCap) %>%
  filter(Age >= 16 & Age <= 17) %>%
  colMeans()
# lung cap is 11.26
df %>%
  select(Age, LungCap) %>%
  filter(Age > 18) %>%
  colMeans()

```
## 1.e 

(lung cap for smokers and non smokers broken into age groups)

We notice a clear trend that smokers have a lower lung capacity compared to non-smokers

```{r}
df %>%
  select(Age, LungCap, Smoke) %>%
  group_by(Smoke) %>%
  filter(Age >= 13) %>%
  summarize_at(vars(LungCap), list(mean = mean))

df %>%
  select(Age, LungCap, Smoke) %>%
  group_by(Smoke) %>%
  filter(Age >= 14 & Age <= 15) %>%
  summarize_at(vars(LungCap), list(mean = mean))

df %>%
  select(Age, LungCap, Smoke) %>%
  group_by(Smoke) %>%
  filter(Age >= 16 & Age <= 17) %>%
  summarize_at(vars(LungCap), list(mean = mean))

df %>%
  select(Age, LungCap, Smoke) %>%
  group_by(Smoke) %>%
  filter(Age > 18) %>%
  summarize_at(vars(LungCap), list(mean = mean))
```
## 1.f 

(correlation and covariance between lung capacity and age)

correlation is at .819 meaning they have a positive correlation of about 82%. This means that there is a connection between the two and when one goes up so does the other. 

```{r}
cov(df$LungCap, df$Age)
cor(df$LungCap, df$Age)
```
# Question 2

```{r}

# creating the Tibble
df <- tibble(X=c(0,1,2,3,4), Freq=c(128,434,160,64,26))

# Creating the probability of an event occurring
df1 <- df %>%
  select(X, Freq) %>%
  mutate(Probability = Freq/sum(Freq))
df1

```

## 2.a

probability of exactly 2 convictions
probability = 19.7%

```{r}

df1 %>%
  select(X, Freq, Probability) %>%
  filter(X == 2)
```

## 2.b

probability of fewer than 2 convictions
probability = 69.2%

```{r}

sum(df1$Probability[1:2])
```

## 2.c

Probability of having 2 or fewer convictions
probability = 88.9%

```{r}

sum(df1$Probability[1:3])
```

## 2.d

probability of having more than 2 convictions
probability = 11.08%

```{r}

sum(df1$Probability[4:5])

```

## 2.e

What is the expected value
expected value is 1.29 convictions 

```{r}

df1 %>%
  select(X, Freq, Probability) %>%
  mutate(expected_value = (0*0.15763547)+(1*0.53448276)+(2*0.19704433)+(3*0.07881773)+(4*0.03201970))
```

## 2.f

What is the variance and standard deviation of the prior convictions
Variance = 25810.8
standard deviation = 160.6574
```{r}
var(df$Freq)
sd(df$Freq)
```

