---
title: "HW1_RoyYoon"
author: "Roy Yoon"
editor: visual
desription: "First homework"
date: "12/20/22"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - Roy Yoon
  - HW
---

```{r}
library(readxl)
library(ggplot2)
```

1. Use the LungCapData to answer the following questions. (Hint: Using dplyr, especially
group_by() and summarize() can help you answer the following questions relatively
efficiently.)

```{r}

lung<- read_xls("~/Documents/603/603_Fall_2022/posts/_data/LungCapData.xls")
lung
```


a. What does the distribution of LungCap look like? (Hint: Plot a histogram with
probability density on the y axis)

```{r}
hist(lung$LungCap)
```
The data presents virtually as a normal distribution 

b. Compare the probability distribution of the LungCap with respect to Males and
Females? (Hint: make boxplots separated by gender using the boxplot() function)

```{r}
#Box plot

ggplot(lung, aes(Gender,LungCap)) +
  geom_boxplot()
```
Male mean is slightly higher than the female mean. However, the means can statisticadlly potentially intersect

c. Compare the mean lung capacities for smokers and non-smokers. Does it make sense?

```{r}
lung %>% 
  group_by(Smoke) %>% 
  summarise(average_lung = mean(LungCap))
```

Matters of the heart are not black and white. However, it is confusing to understand that in this data smokers have a higher lung capcity. I wonder about moderating and mediating effects.

d. Examine the relationship between Smoking and Lung Capacity within age
groups: “less than or equal to 13”, “14 to 15”, “16 to 17”, and “greater than or equal to 18”.

```{r}
#categorical variable of age_groups 
a<- lung %>% 
group_by(Smoke, LungCap) %>% 
  summarise(age_group = case_when(Age<=13 ~ "13 & under",
                                  Age ==14 | Age == 15 ~ "14 to 15",
                                  Age== 16 | Age == 17 ~ "16 to 17",
                                  Age>=18 ~ "18 & older")) 
a


#mean of lung capacity with new variable
a %>% 
  group_by(Smoke, age_group) %>% 
  summarise(avg_lung_cap = mean(LungCap)) %>% 
  arrange(desc(avg_lung_cap))


#histogram
ggplot(a, aes(x = LungCap)) + facet_grid(Smoke ~age_group) +geom_histogram()
```


e. Compare the lung capacities for smokers and non-smokers within each age
group. Is your answer different from the one in part c. What could possibly be going on here?

```{r}
# Mean of non-smokers 13 & younger
a %>% 
  filter(Smoke == 'no' & age_group == '13 & under') %>% 
  pull(LungCap) %>% 
  mean()

# Mean of smokers 13 & younger
a %>% 
  filter(Smoke == 'yes' & age_group == '13 &under') %>% 
  pull(LungCap) %>% 
  mean()

#Mean of non-smokers 14 to 15
a %>% 
  filter(Smoke == 'no' & age_group == '14 to 15') %>% 
  pull(LungCap) %>% 
  mean()

#Mean of smokers 14 to 15
a %>% 
  filter(Smoke == 'yes' & age_group == '14 to 15') %>% 
  pull(LungCap) %>% 
  mean()

#Mean of non-smokers 16 to 17
a %>% 
  filter(Smoke == 'no' & age_group == '16 to 17') %>% 
  pull(LungCap) %>% 
  mean()

#Mean of smokers 16 to 17
a %>% 
  filter(Smoke == 'yes' & age_group == '16 to 17') %>% 
  pull(LungCap) %>% 
  mean()

 #Mean of non-smokers 18 & older
a %>% 
  filter(Smoke == 'no' & age_group == '18 & older') %>% 
  pull(LungCap) %>% 
  mean()

# Mean of smokers 18 & older
a %>% 
  filter(Smoke == 'yes' & age_group == '18 & older') %>% 
  pull(LungCap) %>% 
  mean()
```


f. Calculate the correlation and covariance between Lung Capacity and Age. (use
the cov() and cor() functions in R). Interpret your results.
```{r}
#Correlation
cor(lung$LungCap, lung$Age)

#Covariance
cov(lung$LungCap, lung$Age)
```

A correlation greater than.7 denotes a sttong correlation between LungCap and Age. A high covariance indicates that there is a positive relationship with LungCap and Age.

Question 2

```{r}
# x= prior convictions 
x<-c(0, 1, 2, 3, 4)
frequency<-c(128, 434, 160, 64, 24)
prison<- tibble(x,frequency)

prison
```


What is the probability that a randomly selected inmate has exactly 2 prior convictions?


```{r}
160/810
```
0.19 probability that a randomly selected inmate has fewer than 2 priorconvictions

b. What is the probability that a randomly selected inmate has fewer than 2 priorconvictions?

```{r}
(128+434)/810
```

0.69 that a randomly selected inmate has fewer than 2 priorconvictions

c. What is the probability that a randomly selected inmate has 2 or fewer prior convictions?

```{r}
(128+434+160)/810
```
0.89 that a randomly selected inmate has 2 or fewer prior convictions

d. What is the probability that a randomly selected inmate has more than 2 prior convictions?

```{r}
(64+24)/810
```
0.10 that a randomly selected inmate has more than 2 prior convictions

e. What is the expected value for the number of prior convictions?

```{r}

g<-((128*0/810) +(434*1/810) +(160*2/810) +(64*3/810) +(24*4/810)) 
mean(g)
```
the expected value for the number of prior convictions is 1.28

f. Calculate the variance and the standard deviation for the Prior Convictions.

```{r}

v<- ((0-1.28)^2) *(128/810) +((1-1.28)^2) * (434/810)+((2-1.28)^2) * (160/810)+((3-1.28)^2) *(64/810) +((4-1.28)^2) * (24/810)

v

sd<- sqrt(v)
sd
```
The variance is 0.856 and the standard deviation is 0.925.
