---
title: "Homework 2"
author: "Saaradhaa M"
date: "10/10/2022"
format: 
  html:
    df-print: paged
    toc: true
    code-copy: true
    code-tools: true
    css: "styles.css"
categories:
  - hw2
---

# Qn 1

```{r}
# calculate standard errors.
SE_B <- 10/sqrt(539)
SE_A <- 9/sqrt(847)

# calculate t-values.
CL <- 0.90  
TA <- (1-CL)/2
tvalue_b <- qt(p = 1-TA, df = 539-1)
tvalue_a <- qt(p = 1-TA, df = 847-1)

# calculate CI for bypass.
CIB <- c(19 - tvalue_b * SE_B,
        19 + tvalue_b * SE_B)
CIB

# calculate CI range for bypass.
(19 + tvalue_b * SE_B) - (19 - tvalue_b * SE_B)
        
# calculate CI for angiography.
CIA <- c(18 - tvalue_a * SE_A,
        18 + tvalue_a * SE_A)
CIA

# calculate CI range for angiography.
(18 + tvalue_a * SE_A) - (18 - tvalue_a * SE_A)
```

The 90% CI is narrower for angiography.

# Qn 2

```{r}
set.seed(0)
prop <- prop.test(x=567, n=1031)
prop
```

The 95% CI is \[`r prop$conf.int`\], which includes the point estimate `r prop$estimate` and excludes 0.5. Hence, we can reject the null hypothesis that the true probability is 0.5 at the 5% significance level, *p* = `r prop$p.value`.

# Qn 3

```{r}
# calculate population SD.
PSD <- (200-30)/4

# calculate sample size.
n <- round(((1.96*PSD)/5)^2)
```

The minimum required sample size is `r n`.

# Qn 4a

Assumptions: H~0~ is true, observations are independent of one another, *y* is continuous and sample is approximately normally distributed. H~0~: μ = 500 H~a~: μ ≠ 500

```{r}
# calculate t-statistic.
t <- (410-500)/(90/sqrt(9))

# calculate p-value.
p <- 2*pt(q=abs(t), df=8, lower.tail=FALSE)
p
```

We can reject the null hypothesis at the 5% significance level, *t*(8) = `r abs(t)`, *p* = `r p`. Female employees' mean income significantly differs from \$500 per week.

\[I have a question - I am confused on whether I was right to use the absolute value here, and when we should use absolute values.\]

# Qn 4b

```{r}
# calculate p-value.
p2 <- pt(q=t, df=8, lower.tail=TRUE)
p2
```

We can reject the null hypothesis at the 5% significance level, *t*(8)= -3, *p* = `r p2`. Female employees' mean income is significantly less than \$500 per week.

# Qn 4c

```{r}
# calculate p-value.
p3 <- pt(q=t, df=8, lower.tail=FALSE)
p3
```

We fail to reject the null hypothesis at the 5% significance level, *t*(8)= -3, *p* = `r p3`. Female employees' mean income is [**not**]{.underline} significantly more than \$500 per week.

# Qn 5a

```{r}
# calculate SD for Jones and Smith.
SD <- 10*sqrt(1000)

# calculate t for Jones.
t_j <- ((519.5-500)/SD) * sqrt(1000)
t_j

# calculate p-value for Jones.
p_j <- 2*(pt(q=t_j, df=999, lower.tail=FALSE))
p_j

# calculate t for Smith.
t_s <- ((519.7-500)/SD) * sqrt(1000)
t_s

# calculate p-value for Smith.
p_s <- 2*(pt(q=t_s, df=999, lower.tail=FALSE))
p_s
```

# Qn 5b

The result is statistically significant for Smith, but not Jones.

# Qn 5c

It is useful to report the exact p-value in cases like this, when the p-value is very close to alpha. It helps the reader to understand (1) why it was/was not rejected, and (2) how much evidence there is against the null hypothesis.

# Qn 6

```{r}
#create variable.
gas_taxes <- c(51.27, 47.43, 38.89, 41.95, 28.61, 41.29, 52.19, 49.48, 35.02, 48.13, 39.28, 54.41, 41.66, 30.28, 18.49, 38.72, 33.41, 45.02)

# do t-test.
tax <- t.test(gas_taxes, alternative="less",mu=45)
tax
```

The 95% CI is \[`r tax$conf.int`\], which includes the estimated mean `r tax$estimate` and excludes 45. Hence, we can reject the null hypothesis at the 5% significance level, *t*(`r tax$parameter`)= `r tax$statistic`, *p* = `r tax$p.value`. The average tax per gallon in the US in 2005 was significantly less than 45 cents.
