---
title: "Homework 2"
author: "Shoshana Buck"
desription: "Second homework on confidence intervals"
date: "10/17/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - Shoshana Buck
  - dataset
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

Construct the 90% confidence interval to estimate the actual mean wait time for each of the two procedures. Is the confidence interval narrower for angiography or bypass surgery?

## Tail area and standard error for Bypass

```{r}

tail_area<- (1-.90)/2
tail_area

standard_error<- 10/sqrt(539)
standard_error
```

## t-value for Bypass

```{r}
t_score<- qt(p= 1-tail_area, df= 538)
t_score
```

## Confidence interval and margin of error for Bypass

```{r}
CI<- c(19 - t_score * standard_error, 19 + t_score * standard_error)
CI

MOE<- t_score *standard_error
MOE *1.41
```

## Standard error of angiography

```{r}
standard_error2<- 9/sqrt(847)
standard_error2
```

## t-score for angiography

```{r}
t_score2<- qt(p= 1-.05, df= 846)
t_score2
```

## Confidence interval and margin of error for angiography

```{r}

CI<- c(18 - t_score2 * standard_error2, 18 + t_score2 * standard_error2)
CI


MOE2<- t_score2 *standard_error2
MOE2 *1.01
```

The Bypass points are \[18.29029 & 19.70971\] days and has a margin of error of +/- 0.7. Whereas the angiography is \[17.49 & 18.50\] days with a margin of error of +/- 0.5. Angigography is more narrower because it has a larger sample size and the range between the high and low end of the confidence interval is smaller.

# Question 2

## Point estimate P

```{r}
s_size<- 1031
b<- 567

point_estimate<- b/s_size
point_estimate
```

## 95% confidence interval for P

```{r}
prop.test(b,s_size)
```

Based off the point estimate, 54% of the adult Americans that were surveyed by the National Center for Public Policy believe that college education is essential for success. 95% confidence interval of adult Americans who believe that college education is essential for success is \[0.5189682 0.5805580\] which contains the true population mean.

# Question 3

Assuming the significance level to be 5%, what should be the size sample?

## Standard Deviation

```{r}
sd<- (200-30)/4
sd
```

## Solving for n

```{r}
#steps for the equation
#1. 5 = 1.96 * (42.5/sqrt(n))

#2. 5 = 8.3/sqrt(n)

#3. 5*sqrt(n)= 83.3

#4. sqrt(n) = 83.3/5

#5. n= (83.3/5)^2

#6. n= 278.89
```

The standard deviation from the data is 42.5. Since we have solved for the standard deviation we can plug it into the CI equation and solve for n.

## Question 4

### A

Test whether the mean income of female employees differs from \$500 per week. Include assumptions, hypotheses, test statistic, and P-value. Interpret the result. \## Assumptions

We are assuming there is normal distribution, the null hypothesis is: μ= 500 and the alternative hypothesis is 500\> μ \<500.

## Standard error

```{r}
s_sizef<- 9
sd<-90
s_meanf<- 410
null_hypo_mean<- 500

standard_errorf<- sd/sqrt(s_sizef)
standard_errorf

```

## t-score

```{r}
t_stat<- (s_meanf-null_hypo_mean)/standard_errorf
t_stat
```

I took the sample mean of 410 subtracted that from the mu = 500 and then divided it by the standard error = 30.

## p-value

```{r}
p_value<- (pt(t_stat, df=8)) *2
p_value
```

The p-value than the 5% significance level so we can reject the null hypothesis in favor of the alternative hypothesis.

## B +C

Report the P-value for Ha : μ \< 500. Interpret. Report and interpret the P-value for H a: μ \> 500.

```{r}
upper_p_value<- (pt(t_stat, df=8, lower.tail = FALSE))
upper_p_value

lower_p_value<- (pt(t_stat, df=8, lower.tail = TRUE))
lower_p_value
```

The upper-tailed p-value is 0.99 and the lower-tailed p-value is 0.008. If you add the two tails together they will equal 1.

## Question 5

Show that t = 1.95 and P-value = 0.051 for Jones. Show that t = 1.97 and P-value = 0.049 for Smith.

```{r}
jones_sample_mean<- 519.5
smith_sample_mean<-519.7
null_hyp<- 500
jones_se<- 10
smith_se<- 10


```

### A: Jones t-score and p-value

```{r}
jones_t_stat<- (jones_sample_mean-null_hyp)/jones_se
jones_t_stat

jones_p_value<- pt(jones_t_stat, df=999, lower.tail = FALSE) *2
jones_p_value
```

### A: Smith t-score and p-value

```{r}
smith_t_stat<-(smith_sample_mean-null_hyp)/smith_se
smith_t_stat

smith_p_value<- pt(smith_t_stat, df=999, lower.tail = FALSE)*2
smith_p_value
```

### B

Using α = 0.05, for each study indicate whether the result is "statistically significant."

The results are "statistically significant when the p-value is smaller than the 0.05. Jones p-value is 0.051 which is greater than the 0.05 significance level which means it is not statistically significant and we cannot reject the null hypothesis. Smith's p-value is 0.49 which is smaller than the significance level which means it is statistically significant and that we can reject the null hypothesis in favor of the alternative hypothesis.

### C

"P ≤ 0.05" versus "P \> 0.05," or as "reject H0" versus "Do not reject H0 ," is a misleading statement without providing the p-values because it makes it seem that there is a drastic difference between Jones and Smith that caused one hypothesis to be statistically significant and the other one not to be. However, when looking at the actual p-value it can be noted that there is a very small difference between the values.

## Question 6

```{r}
gas_taxes <- c(51.27, 47.43, 38.89, 41.95, 28.61, 41.29, 52.19, 49.48, 35.02, 48.13, 39.28, 54.41, 41.66, 30.28, 18.49, 38.72, 33.41, 45.02)
```

```{r}

t.test(gas_taxes, mu=45, alternative = 'less')
```

Is there enough evidence to conclude at a 95% confidence level that the average tax per gallon of gas in the US in 2005 was less than 45 cents?

At the 95% confidence level the p-value is 0.03 which is less than the 5% significance level. This proves that we can reject the null hypothesis and that the average tax per gallon of gas in the US in 2005 was less than 45 cents.
