---
title: "Homework 3"
author: "Saaradhaa M"
date: "10/28/2022"
format: 
  html:
    df-print: paged
    toc: true
    code-copy: true
    code-tools: true
    css: "styles.css"
categories:
  - hw3
---

# Qn 1.1.1

The predictor is `ppgdp` and the response is `fertility`.

# Qn 1.1.2

```{r}
# load libraries.
library(tidyverse)
library(alr4)
library(smss)

# load dataset.
data(UN11)

# draw scatterplot.
scatterplot(fertility ~ ppgdp, UN11)
```

No, the graph seems curvilinear.

# Qn 1.1.3

```{r}
# draw scatterplot.
scatterplot (log(fertility) ~ log(ppgdp), UN11)
```

Yes, the simple linear regression model now seems plausible.

# Qn 2a

We can test this using the UN11 dataset since `ppgdp` is in US dollars.

```{r}
# create new variable.
UN11$british <- 1.33*UN11$ppgdp

# check slope.
summary(lm(fertility ~ british, UN11))
summary(lm(fertility ~ ppgdp, UN11))
```

The magnitude of the slope has reduced very slightly, although adjusted R\^2 has not.

# Qn 2b

We can test this too.

```{r}
# correlation with US dollars.
cor(UN11$ppgdp, UN11$fertility)

# correlation with British pounds.
cor(UN11$british, UN11$fertility)
```

Since we multiplied by a constant, the correlation remains the same.

# Qn 3

```{r}
# load dataset.
data(water)

# generate scatterplots.
pairs(water)
```

Stream runoff (BSAAM) seems to have a positive linear relationship with precipitation at OPSLAKE, OPRC and OPBPC; but not with precipitation at APMAM, APSAB or APSLAKE. Stream runoff also seems to be fairly constant (?) over the years.

# Qn 4

```{r}
# load dataset.
data(Rateprof)

# create subset.
rateprof <- Rateprof %>% select(quality, helpfulness, clarity, easiness, raterInterest)

# generate scatterplots.
pairs(rateprof)
```

Quality, helpfulness and clarity have the clearest linear relationships with one another. Easiness and raterInterest do not seem to have linear relationships with the other variables.

# Qn 5a

```{r}
# load dataset.
data(student.survey)
glimpse(student.survey)

# generate plots.
boxplot(pi ~ re, student.survey)
scatterplot(hi ~ tv, student.survey)
```

-   Religiosity and conservatism seem to have a positive relationship.
-   High school GPA and TV-watching seem to have a negative relationship.

# Qn 5b

```{r}
# change pi to numeric variable.
student.survey$pi <- as.numeric(student.survey$pi)

# removing ordering in re and rename it.
levels(student.survey$re) <- c("N", "O", "M", "E")
student.survey$re <- factor(student.survey$re, ordered = FALSE)

# run regression models.
summary(lm(pi ~ re, student.survey))
summary(lm(hi ~ tv, student.survey))
```

-   Those who attended religious services most weeks/every week were significantly more likely to be conservative than those who never did, *p* \< .001. There was no significant difference in political ideology between those who occasionally attended religious services and those who never did.
-   Watching less hours of TV per week was associated with higher high-school GPAs, *p* \< .05. That being said, as the R^2^ is fairly low, hours of TV watching is not a great predictor of high school GPA.
