---
title: "Homework 3"
author: "Roy Yoon"
desription: "HW3"
date: "10/31/22"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3
  - Roy Yoon
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(alr4)
library(smss)
library(ggplot2)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

**United Nations** (Data file: UN11in alr4) The data in the file UN11 contains several variables, including ppgdp, the gross national product per person in U.S. dollars, and fertility, the birth rate per 1000 females, both from the year 2009. The data are for 199 localities, mostly UN member countries, but also other areas such as Hong Kong that are not independent countries. The data were collected from the United Nations (2011). We will study the dependence of fertility on ppgdp.

## 1.1.1

Identify the predictor and the response.

Predictor: gross national product per person in U.S. dollars (ppdgp)

Response: fertility

```{r UN11 data}
UN11
```

## 1.1.2

Draw the scatterplot of fertility on the vertical axis versus ppgdp on the horizontal axis and summarize the information in this graph. Does a straight-line mean function seem to be plausible for a summary of this graph?

```{r select fertility and ppgdp}

fertiity_ppgddp <- UN11 %>%
  select(fertility,ppgdp )

fertiity_ppgddp

```

```{r fertility and ppgdp ggplot}
fertility_ppgdp_plot <- ggplot(data = fertiity_ppgddp, aes(x= ppgdp, y = fertility)) + geom_point()

fertility_ppgdp_plot
```

A straight line mean function does not seem plausible for a summary of this graph as a there does not seem to be a plausible predictive straight line that would be able to describe the tendencies in the scatter plot. Rather the shape seems to be asymptotic.

## 1.1.3

Draw the scatterplot of log(fertility) versus log(ppgdp) using natural logarithms. Does the simple linear regression model seem plausible for a summary of this graph? If you use a different base of logarithms, the shape of the graph won't change, but the values on the axes will change.

```{r ggplot log ppgdp}

fertility_logppgdp_plot <- ggplot(data = fertiity_ppgddp, aes(x= log(ppgdp), y = fertility)) + geom_point()

fertility_logppgdp_plot

```

A simple linear regression model seems plausible for a summary of this graph as a there seems to be a plausible predictive straight line(negative slope) that would be able to describe the tendencies in the scatter plot of log(ppgdp)

```{r ggplot log fertility}
logfertility_ppgdp_plot <- ggplot(data = fertiity_ppgddp, aes(x= ppgdp, y = log(fertility))) + geom_point()

logfertility_ppgdp_plot

```

A simple linear regression model does not seem plausible for a summary of this graph as a there does not seem to be a plausible predictive straight line that would be able to describe the tendencies in the scatter plot of log(fertility)

# Question 2

Annual income, in dollars, is an explanatory variable in a regression analysis. For a British version of the report on the analysis, all responses are converted to British pounds sterling (1 pound equals about 1.33 dollars, as of 2016).

## A

How, if at all, does the slope of the prediction equation change?

suppose there is 10, 20, and 30 British pounds sterling and 13.3, 26.6, and 39.9 USD.

The difference between the 10, 20, and 30 British pounds sterling is 10 as the values increment.

The difference between 13.3, 26.6, and 39.9 USD is 13.3 as the values increment.

13.3 is a greater rate of change than 10. Thus looking at the slope as the rate of change, there is a difference in the rate of change.

## B

How, if at all, does the correlation change?

Correlation is standardized version of the slope. The unit of measurement does not affect the correlation value.

There will be no change to the correlation. The correlation observes how close the dad points are to the line of the data. If all points are converted to the same scale, then how close the points remain to the line should remain constant.

# Question 3

**Water runoff in the Sierras** (Data file: water in alr4) Can Southern California's water supply in future years be predicted from past data? One factor affecting water availability is stream runoff. If runoff could be predicted, engineers, planners, and policy makers could do their jobs more efficiently. The data file contains 43 years' worth of precipitation measurements taken at six sites in the Sierra Nevada mountains (labeled APMAM, APSAB, APSLAKE, OPBPC, OPRC, and OPSLAKE) and stream runoff volume at a site near Bishop, California, labeled BSAAM. **Draw the scatterplot matrix for these data and summarize the information available from these plots. (Hint: Use the pairs() function.)**

```{r}
pairs(water)

```

There seems to be a positive correlation between BSAM(stream runoff volume) and OPSLAKE, OPRC, and OPBP (different precipitation measurements).

There seems to be less of an identifiable correlation between BSAM and APSLAKE, APSAB, APMAM. The data seems more scattered at various points with some being more bunched near the lower x and y points.


# Question 4

**Professor ratings** (Data file: Rateprof in alr4) In the website and online forum RateMyProfessors.com, students rate and comment on their instructors. Launched in 1999, the site includes millions of ratings on thousands of instructors. The data file includes the summaries of the ratings of 364 instructors at a large campus in the Midwest (Bleske-Rechek and Fritsch, 2011). Each instructor included in the data had at least 10 ratings over a several year period. Students provided ratings of 1--5 on quality, helpfulness, clarity, easiness of instructor's courses, and raterInterest in the subject matter covered in the instructor's courses. The data file provides the averages of these five ratings. **Create a scatterplot matrix of these five variables. Provide a brief description of the relationships between the five ratings**.

```{r}
Rateprof

RateMyProf <- Rateprof %>% 
  select(quality, helpfulness, clarity, easiness, raterInterest)

RateMyProf

pairs(RateMyProf)
```

There seems to be a more clear positive correlation between(x,y): quality and helpfulness, quality and clarity, helpfulness and quality, helpfulness and clarity, clarity and helpfulness, clarity and quality. It seems that easiness and raterInterest, whether observed as a explanatory or response variable, seems to yield constant results where the data points are scattered or gathered in similar points with less of a clearly identifiable correlation.


# Question 5

For the student.survey data file in the smss package, conduct regression analyses relating (by convention, y denotes the outcome variable, x denotes the explanatory variable) (i) y = political ideology and x = religiosity, (ii) y = high school GPA and x = hours of TV watching.*You can use ?student.survey in the R console, after loading the package, to see what each variable means.*

## A

Graphically portray how the explanatory variable relates to the outcome variable in each of the two cases.

### (i) y = political ideology and x = religiosity

```{r}
student.survey


a <- ggplot(data = student.survey, aes(x= as.numeric(pi), y = as.numeric(re))) + geom_smooth()

a

```

### (ii) y = high school GPA and x = hours of TV watching

```{r}

b <- ggplot(data = student.survey, aes(x= as.numeric(tv), y = as.numeric(hi))) + geom_smooth()

b
```

## B

Summarize and interpret results of inferential analyses.

(i) y = political ideology and x = religiosity

(ii) y = high school GPA and x = hours of TV watching

the relationship between the average number of hours per week that one watches tv and the high school gpa on a four point square is negative. The more hours of tvs one watches per week, the lower the gpa is.


```{r}
summary(lm(hi ~ tv, data = student.survey))
```

