---
title: "Homework 3"
author: "Yakub Rabiutheen"
description: "The first homework on descriptive statistics and probability"
date: "09/20/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Question 1

## a

The data in the file UN11 contains several variables, including ppgdp, the gross national product per person in U.S. dollars, and fertility, the birth rate per 1000 females, both from the year 2009. The data are for 199 localities, mostly UN member countries, but also other areas such as Hong Kong that are not independent countries. The data were collected from the United Nations (2011). We will study the dependence of fertility on ppgdp.



```{r, echo=T}
library(tidyverse)
library(alr4)
library(smss)
```




```{r, echo=T}
##load data
data(UN11)
```

# Qn 1.1.1

The predictor is `ppgdp` and the response is `fertility`.




```{r}
# Qn 1.1.1 Standard Scatterplot
library(ggplot2)
ggplot(data = UN11, aes(x=ppgdp,y=fertility)) + geom_point()
```





# Qn 1.1.2

```{r}
# Log scatterplot.
scatterplot (log(fertility) ~ log(ppgdp), UN11)
```





# Question-2.a


Annual income, in dollars, is an explanatory variable in a regression analysis. For a British version of the report on the analysis, all responses are converted to British pounds sterling (1 pound equals about 1.33 dollars, as of 2016).

```{r}
usdollar<- (1:10)
pound<- seq(1.33,13.3, length.out = 10)

slope<-(usdollar/pound)
slope
```

To convert from USD to GBP, the value of the response must be divided by 1.33. Same goes for the slope.

## Question-2.b

How, if at all, does the correlation change?
```{r}
cor.test(usdollar,pound)
```

Currency Changes do not affect correlation.

## Question-3


Water runoff in the Sierras (Data file: water in alr4) Can Southern California's water supply in future years be predicted from past data? One factor affecting water availability is stream runoff. If runoff could be predicted, engineers, planners, and policy makers could do their jobs more efficiently. The data file contains 43 years' worth of precipitation measurements taken at six sites in the Sierra Nevada mountains (labeled APMAM, APSAB, APSLAKE, OPBPC, OPRC, and OPSLAKE) and stream runoff volume at a site near Bishop, California, labeled BSAAM. Draw the scatterplot matrix for these data and summarize the information available from these plots. (Hint: Use the pairs() function.)






```{r}
pairs(water_supply,main = "Sierra Southern California Water Supply Runoff",
      pch = 21, bg = "green")
```


```{r}
lm_water_supply<-lm(BSAAM~APMAM+APSAB+APSLAKE+OPBPC+OPRC+OPSLAKE,data = water)
summary(lm_water_supply)
```
The following variables  OPBPC, OPRC, OPSLAKE are correlated with each other.



## Q.4

```{r}
data("Rateprof")
pairs(~Rateprof$quality+Rateprof$helpfulness+Rateprof$clarity+Rateprof$easiness+Rateprof$raterInterest, lwd=2, labels = c("QUALITY", "HELPFULNESS", "CLARITY", "EASINESS", "Rater INTEREST"), pch=19, cex = 0.75, col = "blue")
```
The following variables   “quality”, “clarity”, and “helpfulnes  are correlated with each other.




## Q.5


```{r}
##load data
data(student.survey)
pi_conv <- as.numeric(student.survey$pi)
re_conv <- as.numeric(student.survey$re)
##run regression analysis
model1 <- lm(pi_conv ~ re_conv, data = student.survey)
summary(model1)
```


```{r}
##run regression analysis
model2 <- lm(hi ~ tv, data = student.survey)
summary(model2)
```


```{r}
library(smss)
data("student.survey")
ggplot(data=student.survey,aes(x=re,fill=pi))+
  geom_bar() + labs(x="Religiosity", fill ="Political Ideology")
```

As shown in the graph,there is a strong correlation association between religiosity and Political Idealogy.

```{r}
data("student.survey")
ggplot(data=student.survey,aes(x=hi, y=tv)) +
  geom_point() + labs(x="High School GPA", y="Hours Watching TV")  
```


There is very little relationship betweeh watching TV and High School GPA.




```{r}
summary(student.survey[,c('pi', 're', 'hi', 'tv')])
```

