---
title: "Homework 3"
author: "Nick Boonstra"
date: "`r Sys.Date()`"
---

```{r setup}

library(tidyverse)
library(alr4)
library(smss)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Question 1

```{r q1 read in}

un11 <- tibble(UN11)

head(un11,20)
summary(un11)

```

This question involves using United Nations data to "study the dependence of `fertility` on `ppgdp`" (the dependence of births per 1000 females on GDP in U.S. dollars per person).

## 1.1.1

In this model, `ppgdp` would be the predictor variable, and `fertility` would be the response variable.

## 1.1.2

```{r un11 lin_lin}

un11 %>% 
  ggplot(aes(x=ppgdp,y=fertility)) +
  geom_point() +
  geom_smooth(method=lm,se=F)



```

A linear OLS regression of a linear-linear model of the data does not seem to fit the data very well. While the regression does capture the general downward trend of the data, it is plain to see in the above visualization that it is not a good fit, particularly for larger values of the response variable (`ppgdp`).

## 1.1.3

```{r un11 log_log}

un11 %>% 
  ggplot(aes(x=log(ppgdp),y=log(fertility))) +
  geom_point() +
  geom_smooth(method=lm,se=F)

```

This log-log model of the data is much better suited to a linear OLS regression fit.

# Question 2

```{r q2 df}

q2 <- data.frame(
  life_expec = c(77, 80, 67, 59, 91),
  usd = c(70000, 91000, 65000, 45000, 122000)
)

q2 <- q2 %>% 
  mutate(gbp = usd * 1.33)

q2

q2 %>% 
  ggplot(aes(x=usd,y=life_expec)) +
  geom_point() +
  geom_smooth(method=lm,se=F)

q2 %>% 
  ggplot(aes(x=gbp,y=life_expec)) +
  geom_point() +
  geom_smooth(method=lm,se=F)

summary(lm(life_expec ~ usd, data = q2))
summary(lm(life_expec ~ gbp, data = q2))

```

As can be seen by this example data, transforming a dataset between USD and GBP (that is, adding a linear coefficient to a variable) does change the slope coefficient of the prediction equation (corresponding to the ratio of the value of the two currencies), but it does not alter the overall correlation between the two variables.

# Question 3

```{r q3}

water_data <- tibble(water)

head(water_data,10)

pairs(water_data[,2:8])

```

The correlation between stream runoff near Bishop, CA and rainfall at these six sites in the Sierra Nevada mountains can be interpreted from the scatterplots along the bottom row or the right-most column in the scatterplot matrix (they are both the same information). From this scatterplot matrix, it would appear that correlation between Bishop, CA runoff and Sierra Nevada rainfall is strongest for the `OPSLAKE`, `OPRC`, and `OPBPC` variables. In these plots, the data most clearly follow a linear trend, and the data are mostly staying close to a fit line. 


# Question 4

```{r q4}

rateprof <- tibble(Rateprof)
head(rateprof,10)

pairs(rateprof[,8:12])

```

There is a strong positive correlation between `quality`, `helpfulness`, and `clarity`; besides that, the trend seems to be upward for all other pairs, but the correlation is not as strong.


# Question 5

```{r q5 read_in}

data(student.survey)

summary(student.survey)

```

Some of the variables in question for the analysis will need to be recoded to ordinal variables before a proper regression analysis can be performed.

```{r recode}

student_survey <- student.survey %>% 
  mutate(pi_ord = case_when(
    pi == "very liberal" ~ -3,
    pi == "liberal" ~ -2,
    pi == "slightly liberal" ~ -1,
    pi == "moderate" ~ 0,
    pi == "slightly conservative" ~ 1,
    pi == "conservative" ~ 2,
    pi == "very conservative" ~ 3
  )) %>% 
  mutate(re_ord = case_when(
    re == "never" ~ 0,
    re == "occasionally" ~ 1,
    re == "most weeks" ~ 2,
    re == "every week" ~ 3
  ))

```

```{r pi_re}

summary(lm(pi_ord ~ re_ord, data = student_survey))

student_survey %>% 
  ggplot(aes(x=re_ord,y=pi_ord)) +
  geom_point(alpha=0.12) +
  geom_smooth(method=(lm))

```
A one-point ordinal increase in church attendance, based upon this response scale, corresponds to nearly a one-point ordinal increase in conservative ideology, with zero church attendance ("Never") corresponding to a "Liberal" ideology, all to a strong level of statistical significance.


```{r hi_tv}

summary(lm(hi ~ tv, data = student_survey))

student_survey %>% 
  ggplot(aes(x=tv,y=hi)) +
  geom_point(alpha=0.5) +
  geom_smooth(method=(lm))

```
Each additional hour of TV watched corresponds to a roughly 0.02 point decrease in a student's high school GPA. However, there are a few outliers, particularly along the predictor variable, that may skew this result.
