---
title: "Homework 3"
author: "Donny Snyder"
desription: "Homework 3 Submission"
date: "10/17/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3
  - regression
---

```{r}
library(tidyverse)
library(alr4)
library(smss)
library(dplyr)
library(ggplot2)
library(GGally)
```

# Question 1

```{r, echo=T}
data <- UN11

ggplot(data, aes(x = ppgdp, y = fertility)) + geom_point()
ggplot(data, aes(x = log(ppgdp), y = log(fertility))) + geom_point()

```

#Question 1.1
The predictor is ppgdp and the response is fertility.

#Question 1.2
A straight-line mean function does not seem to be plausible for this graph.

#Question 1.3
A simple linear regression model does seem plausible for a summary of the log log graph.

```{r, echo=T}
data$ppgdp2 <- data$ppgdp*0.75

model1 <- lm(fertility ~ ppgdp, data)
summary(model1)

model2 <- lm(fertility ~ ppgdp2, data)
summary(model2)

cor(data$fertility, data$ppgdp)
cor(data$fertility, data$ppgdp2)
```

# Question 2(a)
The slope of the prediction equation will increase, as the units of the explanatory variable are decreased.

# Question 2(b)
The correlation will stay the same.

# Question 3

```{r, echo=T}
watData <- water
pairs(watData)
ggpairs(watData)

```
OPBPC, OPRC, and OPSLAKE all seem to be highly correlated with BSAAM.

# Question 4

```{r, echo=T}
profData <- Rateprof
profData <- data.frame(profData$quality, profData$helpfulness, profData$clarity, profData$easiness, profData$raterInterest)
pairs(profData)

```
It seems as if quality, helpfulness and clarity are all highly interrelated. easiness and raterInterest are not as highly correlated.

#Question 5
```{r, echo=T}
stud <- as.data.frame(student.survey)

stud$piNum <- NA
stud$reNum <- NA

x = 1
while(x <= 60){
  if(stud$pi[x] == "very liberal"){
    stud$piNum[x] = -3
  }
  if(stud$pi[x] == "liberal"){
    stud$piNum[x] = -2
  }
  if(stud$pi[x] == "slightly liberal"){
    stud$piNum[x] = -1
  }
  if(stud$pi[x] == "moderate"){
    stud$piNum[x] = 0
  }
  if(stud$pi[x] == "very conservative"){
    stud$piNum[x] = 3
  }
  if(stud$pi[x] == "conservative"){
    stud$piNum[x] = 2
  }
  if(stud$pi[x] == "slightly liberal"){
    stud$piNum[x] = 1
  }
  
  
  if(stud$re[x] == "never"){
    stud$reNum[x] = 0
  }
  if(stud$re[x] == "occasionally"){
    stud$reNum[x] = 1
  }
  if(stud$re[x] == "most weeks"){
    stud$reNum[x] = 2
  }
  if(stud$re[x] == "every week"){
    stud$reNum[x] = 3
  }
  x = x + 1
}

model3 <- lm(piNum~reNum, stud)
summary(model3)

model4 <- lm(hi~tv, stud)
summary(model4)

ggplot(stud, aes(x = reNum, y = piNum)) + geom_jitter()

ggplot(stud, aes(x = tv, y = hi)) + geom_jitter()

```
It seems like the results are that political ideology tends to be more right-leaning as religiosity increases. As hours of tv watching tends to go down, high school GPA tends to go up. These relationships are both statistically significant.
