---
title: "Homework 4"
author: "Saaradhaa M"
date: "11/14/2022"
format: 
  html:
    df-print: paged
    toc: true
    code-copy: true
    code-tools: true
    css: "styles.css"
categories:
  - hw4
---

```{r}
#| label: setup
#| warning: false
#| message: false
library(tidyverse)
library(alr4)
library(smss)
library(interactions)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Qn 1A

```{r}
#| label: 1A
# predicted price.
yhat <- -10536 + 53.8*1240 + 2.84*18000

# residual.
145000-yhat
```

The model under-predicts by \$37,704.

# Qn 1B

It is predicted to increase by \~\$53.80. This is because the effects of `x1` and `x2` on `y` are independent of one another, and there is no interaction between them.

# Qn 1C

```{r}
#| label: 1C
53.8/2.84
```

It would need to increase by \~19x.

# Qn 2A

```{r}
#| label: 2A
# load dataset.
data(salary)

# run model.
summary(lm(salary ~ sex, salary))
```

The mean salary for men and women does not differ, *p* = 0.07 (i.e., it is the same).

# Qn 2B

```{r}
#| label: 2B
# run model.
model2B <- lm(salary ~ ., salary)
summary(model2B)

# get confidence interval.
confint(model2B)
```

Sex is still not a significant predictor of salary. The 95% CI for difference in salary between males and females is approx. -698 to 3031.

# Qn 2C

`rank` and `year` are significant predictors of salary, while all others were not.

Both `rank` and `year` positively predict salary: Associate Professors and full Professors were likely to earn quite a bit more than Assistant Professors, while professors with more years in their current rank also earned more.

Looking at the magnitude of the coefficients, `rank` has a greater impact on salary than `year` does.

# Qn 2D

```{r}
#| label: 2D
# change baseline category.
salary$rank <- relevel(salary$rank, ref = 'Assoc')

# re-run model.
model2D <- lm(salary ~ ., salary)
summary(model2D)
```

We can see that `rank` now both negatively and positively predicts `salary`. Specifically, Assistant Professors earn less than Associate Professors do, at the same magnitude that was observed in `model1` (approx. \~\$5292).

# Qn 2E

```{r}
#| label: 2E
model2E <- lm(salary ~ . - rank, salary)
summary(model2E)
```

Excluding `rank`, both adjusted R^2^ and the overall F-statistic drop. `degree` and `ysdeg` are now significant as well. Curiously, those with a PhD earn less than those with an MS. As years since highest degree increases, so does salary.

# Qn 2F

```{r}
#| label: 2F1
# create new variable.
salary <- mutate(salary, HireTime = case_when(ysdeg <= 15 ~ "1", ysdeg > 15 ~ "0"))
salary$HireTime <- as.numeric(salary$HireTime)

# run correlation matrix to check for multicollinearity. it's important to do so especially because HireTime is derived from ysdeg, so we would expect them to be highly correlated.
subset <- salary %>% select(year, ysdeg, HireTime)
matrix <- cor(subset, use="complete.obs")
```

I would remove `ysdeg` from the regression model since it is highly correlated with `HireTime`, *r* = 0.84.

```{r}
#| label: 2F2
model2F <- lm(salary ~ . - ysdeg, salary)
summary(model2F)
```

Yes, those hired by the new Dean are making more, *p* = .05.

# Qn 3A

```{r}
#| label: 3A
# load dataset.
data(house.selling.price)

# run model.
model3A <- lm(Price ~ Size + New, house.selling.price)
summary(model3A)
```

Both `Size` and `New` significantly positively predict selling price. As each predictor goes up by 1 unit, selling price rises by \$116 and \$57,736 respectively.

# Qn 3B

```{r}
#| label: 3B
# new homes.
new <- house.selling.price %>% filter(New == 1)
model3B1 <- lm(Price ~ Size, new)
summary(model3B1)

# old homes.
old <- house.selling.price %>% filter(New == 0)
model3B2 <- lm(Price ~ Size, old)
summary(model3B2)
```

Size significantly positively predicts price for both new and old houses, but by a greater magnitude for new houses:

-   New: E(Price) = 166\*Size - 100,755

    -   Adjusted R^2^ for the model is also much higher (0.90 vs. 0.57).

-   Old: E(Price) = 104\*Size - 22,228

# Qn 3C

-   New: E(Price) = 166\*3000 - 100,755 = \$397,245

-   Old: E(Price) = 104\*3000 - 22,228 = \$289,772

# Qn 3D

```{r}
#| label: 3D
model3D <- lm(Price ~ Size*New, house.selling.price)
summary(model3D)
interact_plot(model3D, pred = Size, modx = New)
```

`Size` and `New` positively interact to affect `Price`, such that `New` amplifies the positive relationship between `Size` and `Price`.

# Qn 3E

Referring to `model3D` above:

-   New: -22,228 + 104\*Size - 78,528 + 62\*Size = 166\*Size - 100,756

-   Old: 104\*Size - 22,228

# Qn 3F

-   New: 166\*3000 -100,756 = \$397,244

-   Old: 104\*3000 - 22,228 = \$289,772

# Qn 3G

-   New: 166\*1500 - 100,756 = \$148,244

-   Old: 104\*1500 - 22,228 = \$133,772

As size of home goes up, the difference in predicted selling prices between old and new homes becomes larger.

# Qn 3H

Questions 3F and 3G demonstrate that `size` affects the relationship between `new` and `price`, which reveals some type of dependency between the predictors. Hence, the model with interaction represents the relationships between `size`, `new` and `price` better.
