---
title: "Homework 4"
author: "Caleb Hill"
desription: "The fourth homework on regressions"
date: "11/12/2022"
format:
  html: 
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw4
  - regression
  - prediction
---

# Question 1

First, let's load the relevant libraries and set all the graph themes to minimal.

```{r, results='hide'}
library(readxl)
library(tidyverse)
library(dplyr)
library(alr4)
library(smss)
theme_minimal()
```

## A

The prediction equation for the following three subsections is: ŷ = −10,536 + 53.8x1 + 2.84x.

```{r}
x <- sum(-10536 + (53.8*1240) + (2.84*18000))
y <- 145000

y - x
```

The predicted sale price is \$107,296.That is a difference (residual) of \$37,704.

## B

```{r}
x <- sum(-10536 + (53.8*500) + (2.84*1000))
y <- sum(-10536 + (53.8*501) + (2.84*1000))
y - x
```

For a fixed lot size, the house selling price is predicted to increase \$53.80 per each square-foot increase. This is because we are multiplying the size of the home (in square feet) by \$53.80.

## C

```{r}
sum(53.8/2.84)
```

Lot size would have to increase by almost 19 square feet to have the same impact as a one-square-foot increase in home size.

# Question 2

```{r}
data(salary)
head(salary)
```

## A

```{r}
t.test(salary ~ sex, salary)
```

While there is a difference in salary by `Sex,` the p-value does not meet the threshold for statistical significance (0.05). We cannot reject the null hypothesis.

## B

```{r}
salary$sex <- relevel(salary$sex, ref = 1)
model1 <- lm(salary ~ sex + rank + degree + year + ysdeg, data = salary)
confint(model1, level = 0.95)
```

The 95% CI for the female sex's impact on salary is between -697.82 to 3030.56.

```{r}
salary$sex <- relevel(salary$sex, ref = 2)

model_relevel_sex <- lm(salary ~ sex + rank + degree + year + ysdeg, data = salary)
confint(model_relevel_sex, level = 0.95)
```

The 95% CI for the male sex's impact on salary is between -3030.56 to 697.82.

## C

```{r}
summary(model1)
```

Three variables reach statistical significance (low p-value): `rankAssoc,` `rankProf,` and `year.` `SexFemale` does not reach statistical significance.

`Year` has the lowest estimate and standard error, about 5x to 10x less than the other two variables. The t-value is less than that of `rankProf,` but it is still the second highest.

## D

```{r}
salary$rank <- relevel(salary$rank, ref = 3)

model_relevel_rank <- lm(salary ~ rank + degree + sex + year + ysdeg, data = salary)
summary(model_relevel_rank)
```

We excluded Assoc from the `rank` variable in this relevel and included Asst and Prof. This has shown a positive relationship between `rank` and `salary` for Prof, but not for Asst, and the variables are statistically significant at the 0.001 scale for both of them. The standard error is also lower for both compared Assoc, though not by much.

## E

```{r}
model2 <- lm(salary ~ degree + sex + year + ysdeg, data = salary)
summary(model2)
```

Excluding rank, we now see that `ysdeg` is the best predictor variable, with the lowest estimate score, standard error, and meets the 0.001 p-value threshold to be statistically significant.

## F

```{r}
salary$dean <- ifelse(salary$ysdeg >= '15', "Old Dean",
                  "New Dean")
table(salary$dean)
model3 <- lm(salary ~ dean, data = salary)
summary(model3)
```

It looks like those hired prior to the new Dean do have a statistically significant impact on salary, though at a minor code of 0.05. What's interesting though is the positive relationship, which rejects the belief that the new Dean has been making more generous offers to new hires. However, we should add some control variables, making sure to avoid multicollinearity. Three variables that would impact multicollinearity are `rank,` `ysdeg,` and `year.`

```{r}
model4 <- lm(salary ~ dean + degree + sex, data = salary)
summary(model4)
```

The standard error drops, but everything else remains very similar.

# Question 3

## A

```{r}
data(house.selling.price)

model5 <- lm(Price ~ Size + New, house.selling.price)
summary(model5)
```

`Size` is statistically significant (SS) at the 0.001 level, while `New` is SS at the 0.01 level. `Size` has a small standard error and larger t-value, while `New` has a larger estimate, standard error, and lower t-value. We could relevel to see if Old houses had a better impact on `Price` but that can be observed in other questions.

```{r}
house.selling.price$New <- relevel(factor(house.selling.price$New), ref = 1)

model_relevel_new <- lm(Price ~ Size + New, house.selling.price)
summary(model_relevel_new)
```

## B

The prediction equation for houses is as follows:

New: Predicted Price = -40,230 + 116.13(Size) + 57,736(New).

Old: Predicted Price = -40,230 + 116.13(Size) + -57,736(Old).

## C

```{r}
sum(-40230 + (116.13*3000) + (57736*1))
sum(-40230 + (116.13*3000) + (-57736*1))
```

For a new house, the predicted selling price is \$365,896. For an old house, it's \$250,424.

## D

```{r}
model6 <- lm(Price ~ Taxes + Size + New, house.selling.price)
summary(model6)
```

## E

The predicted selling price for the new model for new homes is as follows:

New: Predicted Price = -21,353 + 37.23(Taxes) + 61.74(Size) + 46,373.70(New)

Old: Predicted Price = -21,353 + 37.23(Taxes) + 61.74(Size) + -46,373.70(Old)

## F

```{r}
sum(-21353 + (37.23*1) + (61.74*3000) + (46373.70*1))
sum(-21353 + (37.23*1) + (61.74*3000) + (-46373.70*1))
```

The predicted selling price for a new 3000 square foot house is \$210,277.90 and for an old house is \$117,530.50.

## G

```{r}
sum(-21353 + (37.23*1) + (61.74*1500) + (46373.70*1))
sum(-21353 + (37.23*1) + (61.74*1500) + (-46373.70*1))
```

The predicted selling price for a new 1500 square foot house is \$117,667.90 and for an old one is \$24,920.53. For each square foot increase, we have a dollar increase of \$61.74.

## H

I prefer the model that includes taxes. This reduces both the residuals and the standard error for the original variables, `Size` and `New.` With a smaller standard error, we should have a more accurate prediction value when attempting to ascertain what the potential sales price for a house is. There is also a larger adjusted R squared percentage. With all that, I would include `Taxes` in the model for the best prediction value.
