---
title: "Homework 5"
author: "Saaradhaa M"
date: "12/06/2022"
format: 
  html:
    df-print: paged
    toc: true
    code-copy: true
    code-tools: true
    css: "styles.css"
categories:
  - hw5
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(alr4)
library(smss)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Qn 1A

`Beds`, as it has the largest p-value.

# Qn 1B

`Size`, as it would produce the smallest p-value.

# Qn 1C

`Beds` is highly correlated with `Size`, so multicollinearity might be a possible reason for this.

# Qn 1D

The model with all predictors has the highest R^2^.

```{r}
#| label: 1d
#| message: false
data(house.selling.price.2)
summary(lm(P ~ ., house.selling.price.2))
```

The model below without `Beds` has the highest adjusted R^2^ and lowest PRESS/BIC/AIC.

```{r}
#| label: 1dar
#| message: false

model <- lm(P ~ .-Be, house.selling.price.2)

PRESS <- function(model) {
  pr <- residuals(model)/(1-lm.influence(model)$hat)
  PRESS <- sum(pr^2)
  return(PRESS)
}

PRESS(model)
BIC(model)
AIC(model)
```

# Qn 1E

I prefer the model without `Beds` because it doesn't seem to be adding great value to the regression model, and the other 4 measures (adjusted R^2^, PRESS, AIC and BIC) are better ways of assessing a multiple regression model than just R^2^.

# Qn 2A

```{r}
data(trees)
model2 <- lm(Volume ~ poly(Girth, 2) + Height, trees)
summary(model2)
```

# Qn 2B

```{r}
par(mfrow = c(2,3)); plot(model2, which = 1:6)
```

Scale-Location plot indicates heteroskedasticity. Cook's distance, residuals vs. leverage and Cook's distance vs. leverage plots indicate 3 outliers. 

# Qn 3A

```{r}
data(florida)
model3a <- lm(Buchanan ~ Bush, florida)
par(mfrow = c(2,3)); plot(model3a, which = 1:6)
```

It seems to stick out in all 6 plots, so I would say it's an outlier.

# Qn 3B

```{r}
model3b <- lm(log(Buchanan) ~ log(Bush), florida)
par(mfrow = c(2,3)); plot(model3b, which = 1:6)
```

Palm Beach still seems to be an outlier, but the diagnostic plots look more normal than in the previous model.