---
title: "Homework 5"
author: "Caleb Hill"
desription: "The fifth homework on regressions"
date: "12/08/2022"
format:
  html: 
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw5
  - regression
---

# Question 1

First, let's load the relevant libraries and set all the graph themes to minimal.

```{r, results='hide'}
library(readxl)
library(tidyverse)
library(dplyr)
library(alr4)
library(smss)
library(stargazer)
```

## A

Variables `Beds` would be removed first. It's P-value does not meet statistically significant threshold.

## B

For the opposite reason, `Size` would be added first.

## C

`Beds` is most likely auto-correlated with `Size`, as bedrooms make up a substantial amount of square footage in a house, driving up the price. However, that is not the only factor if pricing, as is most likely why the number -- which can be high -- does not necessarily mean we see an improved price. Conversely, this is not the case for bathrooms.

## D

```{r}
data(house.selling.price.2)
model_1 <- lm(P ~ S + Ba, data = house.selling.price.2)
plot(model_1)
stargazer(model_1)
```

```{r}
PRESS <- function(model) {
  i <- residuals(model)/(1 - lm.influence(model)$hat)
  sum(i^2)
}

PRESS(model_1)
```

```{r}
broom::glance(model_1)
```

Note: observation #5 is an outlier and violates many of these diagnostic plots. If we log transform either P or S, observation #7 does the same. Therefore, we will keep the model as-is and not transform it.

R2 = 0.83
Adjusted R2 = 0.83
PRESS = 34.174.5
AIC = 809.23
BIC = 819.36

## E

I chose this model as the other two variables `Beds` and `New` either do not meet the P-value threshold or have a low correlation to `Price.` The other two variables have a high correlation with `Price.`

# Question 2

```{r}
data(trees)
head(trees)
```

## A

```{r}
model_q2 <- lm(Volume ~ Girth + Height, data = trees)
```

Fitted.

## B

```{r}
plot(model_q2)
```

Both of the Residuals vs. Fitted plots are violated, as they form a U-shape instead of a horizontal line; Cook's Distance plot is violated as row 31 is outside of the 0.5 dotted line. The Normal Q-Q plot seems fine. The Scale Location plot does not look good, as it has a U-shape as well.

# Question 3

```{r}
data(florida)
head(florida)
```

## A

```{r}
model_q3a <- lm(Buchanan ~ Bush, data = florida)
plot(model_q3a)
```

Yes, Palm Beach County is an outlier. This is especially apparent with Cook's Distance plot, as it is outside the 1.0 dotted line.

## B

```{r}
model_q3b <- lm(log(Buchanan) ~ log(Bush), data = florida)
plot(model_q3b)
```

The findings do change, as all observations that violated the tests are now within the lines/meet assumptions.

