---
title: "Homework 1"
author: "Mani Kanta Gogula"
description: "Homework_1_603"
date: "10/3/2022"
format:
  html:
    df-print: paged
    css: styles.css
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(readxl)
library(ggplot2)
library(stats)

knitr::opts_chunk$set(echo = TRUE)
```

## Question 1


```{r}
Lung_data<- read_excel("C:/Users/manik/Desktop/LungCapData.xls")
Lung_data
```

Given data consists of 725 rows and 6 columns



a. What does the distribution of LungCap look like?

```{r}
Lung_data %>%
  ggplot(aes(LungCap, ..density..)) +
  geom_histogram() +
  geom_density(color = "Red") +
  theme_classic() + 
  labs(title = "Probability distribution of LungCap", x = "Lung Capcity", y = "Probability density")
```

Based on above histogram , we can say the distribution is very close to the normal distribution



Compare the probability distribution of the LungCap with respect to Males and
Females?

```{r}
Lung_data %>%
  ggplot(aes(y = dnorm(LungCap), color = Gender)) +
  geom_boxplot() +
  labs(title = "Probability distribution of LungCap based on gender", y = "Probability density")
```





Compare the mean lung capacities for smokers and non-smokers. Does it make
sense?

```{r}
Mean_smokers <- Lung_data %>%
  group_by(Smoke) %>%
  summarise(mean = mean(LungCap))
Mean_smokers
```

The mean of the lung capacity who smokes is greater than the people who doesnt smoke which doesnt make any sense in practical



Examine the relationship between Smoking and Lung Capacity within age
groups: “less than or equal to 13”, “14 to 15”, “16 to 17”, and “greater than or
equal to 18”.

```{r}
Lung_data <- mutate(Lung_data, AgeGrp = case_when(Age <= 13 ~ "less than or equal to 13",
                                    Age == 14 | Age == 15 ~ "14 to 15",
                                    Age == 16 | Age == 17 ~ "16 to 17",
                                    Age >= 18 ~ "greater than or equal to 18"))

Lung_data %>%
  ggplot(aes(y = LungCap, color = Smoke)) +
  geom_histogram(bins = 25) +
  facet_wrap(vars(AgeGrp)) +
  theme_classic() + coord_flip()
  labs(title = "Relationship of LungCap and Smoke based on age categories", y = "Lung Capacity", x = "Frequency")
  
```
                                    



Compare the lung capacities for smokers and non-smokers within each age
group. Is your answer different from the one in part d. What could possibly be
going on here?

```{r}
Lung_data %>%
  ggplot(aes(x = Age, y = LungCap, color = Smoke)) +
  geom_line() +
  theme_classic() + 
  facet_wrap(vars(Smoke)) +
  labs(title = "Relationship of LungCap and Smoke based on age", y = "Lung Capacity", x = "Age")
```

Form the above data we can compare 1d and 1e and can say the results are pretty similar. Only 10 and above age group smoke.



Calculate the correlation and covariance between Lung Capacity and Age. (use
the cov() and cor() functions in R). Interpret your results.

```{r}
Covariance_LA <- cov(Lung_data$LungCap, Lung_data$Age)
Correlation_LA <- cor(Lung_data$LungCap, Lung_data$Age)
Covariance_LA
Correlation_LA
```

From the above result we can say that both covariance and correlation is positive and which indicates direct relationship that means Lungcapacity increases as age increases

## Question 2



```{r}
Prior_convitions <- c(0:4)
Inmate_count <- c(128, 434, 160, 64, 24)
IP<- data_frame(Prior_convitions, Inmate_count)
IP
```

```{r}
IP <- mutate(IP, Probability = Inmate_count/sum(Inmate_count))
IP
```




What is the probability that a randomly selected inmate has exactly 2 prior
convictions?

```{r}
IP %>%
  filter(Prior_convitions == 2) %>%
  select(Probability)
```



What is the probability that a randomly selected inmate has fewer than 2 prior
convictions?

```{r}
p_2 <- IP %>%
  filter(Prior_convitions < 2)
sum(p_2$Probability)
```



What is the probability that a randomly selected inmate has 2 or fewer prior
convictions?

```{r}
p <- IP %>%
  filter(Prior_convitions <= 2)
sum(p$Probability)
```



What is the probability that a randomly selected inmate has more than 2 prior
convictions?

```{r}
P_3 <- IP %>%
  filter(Prior_convitions > 2)
sum(P_3$Probability)
```



What is the expected value for the number of prior convictions?

```{r}
IP <- mutate(IP, Wm = Prior_convitions*Probability)
expe<- sum(IP$Wm)
expe
```



Calculate the variance and the standard deviation for the Prior Convictions.

```{r}
var_ <-sum(((IP$Prior_convitions-expe)^2)*IP$Probability)
var_
```

standard deviation:

```{r}
sqrt(var_)
```

