---
title: 'Homework #5'
author: "Kalimah Muhammad"
date: "12/9/2022"
desription: "Homework #5"
format:
  
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories: 
  - hw5
  - Kalimah Muhammad
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(dplyr)
library(ggplot2)
library(alr4)
library(smss)
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
data("house.selling.price.2") #load housing data
summary(lm(P ~ ., data = house.selling.price.2))
```

### 1.A

For backward elimination, the first variable that should be deleted is Beds as it has the largest p-value in the regression model at 0.486763.

### 1.B

For forward selection, the first variable that should be added is New since it has the lowest p-value at 4.3e-06.

### 1.C

The BEDS variable could have a large p-value and high correlation because it no longer becomes significant compared to the other variables are added.

### 1.D

```{r}
#model 1
house_model1 <-lm(P ~ New, data = house.selling.price.2)
summary(house_model1)

#model 2
house_model2 <-lm(P ~ S, data = house.selling.price.2)
summary(house_model2)

#model 3
house_model3 <-lm(P ~ Ba, data = house.selling.price.2)
summary(house_model3)

#model 4
house_model4 <-lm(P ~ Be, data = house.selling.price.2)
summary(house_model4)

#summary of AIC comparisons
house.selling.price.2%>%
summarise(AIC(house_model1), AIC(house_model2), AIC(house_model3), AIC(house_model4))
  
#summary of BIC comparisons        
house.selling.price.2%>%
summarise(BIC(house_model1), BIC(house_model2), BIC(house_model3), BIC(house_model4))
```

Using software with these four predictors, find the model that would be selected using each criterion: a. R2 - Model2 with the Size variable would be the best with the highest R-squared value of 0.8079. b. Adjusted R2 - Model2 would be the best fit with the highest value at 0.8058. c. PRESS - Model2 is the best fit. d. AIC - Model2 with the Size would be the best with an AIC of 820.1439. e. BIC - Model2 with the Size would be the best with an BIC of 827.74.

### 1.E

The model using the Size variable resulted in the best fit model.

## Question 2

```{r}
data("trees")
```

### 2.A

```{r}
tree_model1 <-lm(Volume ~ Girth + Height, trees)
summary(tree_model1)

tree_model2 <- lm(Volume ~ Girth * Height, trees)
summary(tree_model2)
```

The best fit model to predict tree volume includes girth plus height as an interaction term. This is displayed in tree_model2.

```{r}
#create matrix of model plots
par(mfrow = c(2,3)); plot(tree_model2, which = 1:6)
```

### 2.B

A few assumptions that may be violated include:\
- Potential heteroskedasticity as suggested in the curved red line in the Scale-Location plot suggesting standardized residuals may be changing as a product of fitting.\
- Observation #18 has significant leverage compared to the observation's (see Cook's distance) and may indicate a larger potential influence on the model.

## Question 3

```{r}
data("florida") #load data
```

### 3.A

Run a simple linear regression model where the Buchanan vote is the outcome and the Bush vote is the explanatory variable. Produce the regression diagnostic plots.

```{r}
fla_model <- (lm(Buchanan ~ Bush, data=florida))
summary(fla_model)
par(mfrow = c(2,3)); plot(fla_model, which = 1:6)
```

Palm Beach County is an outlier. This may be due to this county having the most Buchanan votes (3404).

### 3.B

```{r}
fla_model2 <- (lm(log(Buchanan) ~ log(Bush), data=florida))
summary(fla_model2)
par(mfrow = c(2,3)); plot(fla_model2, which = 1:6)
```

Palm Beach County, remains as the most prominent outlier. However, the second model highlights a second tier of outliers including Glades and Liberty.
