---
title: "HW5"
author: "Ken Docekal"
desription: "Homework 5"
date: "12/10/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw5
  - Ken Docekal

---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(AER)
library(stargazer)

knitr::opts_chunk$set(echo = TRUE)
```



## Q1


```{r}

library(smss) 

data('house.selling.price.2', package = 'smss')

```

```{r}

summary(house.selling.price.2)

tibble(house.selling.price.2)

```


```{r}

cor(house.selling.price.2)

lm(P ~ S + Be + Ba + New, data = house.selling.price.2) |> summary()

```



# A

Using backward elimination Beds would be eliminated first as it is the variable with the largest p-value. 


# B 

If we employ forward selection, as the variables with the lowest p-value, New or Size would be selected first. 
 
 
# C 

Bed's large p-value may be due to its high correlation with Size and therefore not statistically significant on its own as the two variables are multicollinear. 


# D

R2 - Using R squared and backward elimination the model below should be selected as it results in the greatest R squared score. 

```{r}

lm(P ~ S + Be + Ba + New, data = house.selling.price.2) |> summary()

```


Adjusted R2 - The model without Bed included is preferred if considering adjusted R squared in backward elimiation instead. 

```{r}

lm(P ~ S + Ba +  New, data = house.selling.price.2) |> summary()

```


PRESS - Using the PRESS method the model with all variables except Bed is also preferred as that model results in the lowest PRESS score of 27860.05. 

```{r}

library(MPV)

```

```{r}

mod<-lm(P ~ S + Be + Ba + New, data = house.selling.price.2) 

mod1<-lm(P ~ S + Ba + New, data = house.selling.price.2) 

mod2<-lm(P ~ S + New, data = house.selling.price.2) 

mod3<-lm(P ~ S + New, data = house.selling.price.2) 

mod4<-lm(P ~ S, data = house.selling.price.2) 

PRESS(mod)
PRESS(mod1)
PRESS(mod2)
PRESS(mod3)
PRESS(mod4)

```


AIC - The model with all variables but Bed, as the model with the lowest AIC score of 789.14, is preferred.

```{r}

AIC(mod)
AIC(mod1)
AIC(mod2)
AIC(mod3)
AIC(mod4)

```


BIC - The model with all variables but Bed, as the model with the lowest BIC score of 801.80, is preferred.

```{r}

BIC(mod)
BIC(mod1)
BIC(mod2)
BIC(mod3)
BIC(mod4)

```


# E

I prefer the model with all variables except Bed. Tests using adjusted R squared, PRESS, AIC, and BIC as predictors all indicate this model has the greatest comparative predictive power. 



## Q2

```{r}

summary(trees)

tibble(trees)

```



# A

```{r}

lm(Volume ~ Girth + Height, data = trees)|> summary()

```


# B

Based on the residual vs fitted and normal Q-Q plots this regression model does not meet the linearity assumption. Additionally the trend line in the scale-location plot indicates non-constant variance. 

```{r}

fit = lm(Volume ~ Girth + Height, data = trees)

plot(fit, which = 1:6)

```



## Q3

```{r}

library(alr4) 

data('florida', package = 'alr4')

summary(florida)

tibble(florida)

```


# A

Based on the linear regression and plots below, Palm Beach is an outlier. The Cook's distance for Palm Beach is larger than 1 indicating an influential observation. 

```{r}

lm(Buchanan ~ Bush, data = florida)|> summary()

fit2 = lm(Buchanan ~ Bush, data = florida)

plot(fit2, which = 1:6)

```


# B

Using log of both variables instead results a stronger model with a much higer adjusted R squared score of .85 compared to the non-log version's score of .38. This change does not seem to result in a change in the influence of Palm Beach however as indicated by similarities in the Cook's Distance and Scale-Location plots below. 

```{r}

lm(log(Buchanan) ~ log(Bush), data = florida)|> summary()

fit3 = lm(log(Buchanan) ~ log(Bush), data = florida)

plot(fit3, which = 1:6)

```






