---
title: "Homework 2"
author: "Megha joseph"
desription: "Homework 2 Assignment"
date: "10//2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - challenge1
  - my name
  - dataset
  - ggplot2
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

##QUESTION1
```{r}
procedure <- c('Bypass', 'Angiography')
samplesize <- c(539, 847)
meanwait <- c(19, 18)
standev <- c(10, 9)

surgdata <- data.frame(procedure, samplesize, meanwait, standev)

surgdata
     
```
##QUESTION2
```{r}
prop.test(567, 1031, conf.level = .95)

```

##QUESTION3
  
```{r}
stdevBooks <- (200-30)/4
margerrorBooks <- (10/2)
zBooks <- 1.96

stdevBooks^2 * (zBooks/margerrorBooks)^2
```
##QUESTION4
##A

  
```{r}
(410-500)/(90/sqrt(9))
pt(-3, 8)*2 
```
##B

  
```{r}
 pt(-3, 8, lower.tail = TRUE)
```
##C

  
```{r}
 pt(-3, 8, lower.tail = FALSE)
```
##QUESTION5
##A1
  
```{r}
 JonesT <- (519.5-500)/10
JonesT
``` 
##A2

```{r}
 JonesP <- pt(1.95, 999, lower.tail = FALSE)*2
JonesP
``` 
##A3

```{r}

SmithT <- (519.7-500)/10
SmithT
``` 
##A4

```{r}

SmithP <- pt(1.97, 999, lower.tail = FALSE)*2
SmithP
``` 
##B

With an α-level of .05, the p-values that both Jones (P=.051) and Smith (P=.049) found are very close to equivalent. Although Jones’ P-value is slightly greater than α=.05 and Smith’s P-value is slightly less than α=.05, the proximity of the results should yield the same conclusion. Both P-values provide moderate evidence to reject the null hypothesis and indicate that the mean is not equal to 500. If we were to technically interpret the P-values, then Jones’ test would fail to reject the null hypothesis, and Smith’s test would reject the null hypothesis.

##C

If we fail to report the P-value and simply state whether the P-value is less than/equal to or greater than the defined significance level of the test, one cannot determine the strength of the conclusion. For example, a P-value of .009 for a significance level of .05 provides much stronger evidence to reject the null than a P-value of .045, however both values allow for rejection of the null at the significance level .05. In the Jones/Smith example, reporting the results only as “P ≤ 0.05” versus “P > 0.05” will lead to different conclusions about very similar results (rejecting versus failing to reject the null).

##QUESTION6

```{r}
gas_taxes <- c(51.27, 47.43, 38.89, 41.95, 28.61, 41.29, 52.19, 49.48, 35.02, 48.13, 39.28, 54.41, 41.66, 30.28, 18.49, 38.72, 33.41, 45.02)
t.test(gas_taxes, mu = 18.4, conf.level = .95)

``` 


  
  
 
  
  
  


