---
title: "Home Work 3"
author: "Megha Joseph"
date: "10/31/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw3
  - Megha Joseph
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(smss)
library(alr4)

knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
data(UN11) 
UN11

```
## 1.1.1

The predictor variable is ppgdp.
The response variable is fertility.

## 1.1.2

# Build scatterplot
# fertility on vertical axis, ppgdp on horizontal axis

```{r}
plot(x = UN11$ppgdp, y = UN11$fertility, xlab = 'ppgdp', ylab = 'fertility')
```
This graph potrays an intense negative relationship between a country’s gross national product per person and fertility rate at first and then there appears to be a slight change in fertility in relationship to ppgdp moving beyond this point. A straight-line mean function does not seem to be an appropriate measure for summary of this graph.

## 1.1.3

# Create scatterplot
# log(fertility) on vertical axis, log(ppgdp) on horizontal axis

```{r}
plot(x = log(UN11$ppgdp), y = log(UN11$fertility), xlab = 'log(ppgdp)', ylab = 'log(fertility')

```
The simple linear regression seems plausible for summary of this graph. 

## Question 2

a. The slope of the prediction equation would change. It would be the initial version’s slope divided by 1.33 to account for the change in unit to pounds.

b.The correlation does not change, because it standardizes the slope (thus is not impacted by unit of measure).

## Question 3
```{r}
data(water)
head(water)
```

```{r}
#load dataset 
data(water)

#scatterplot matrix
pairs(water)

#calculate the summary
summary(water)
```
In this scatterplot matrix, the precipitation levels for the ‘A’ named lakes seem to have a positive (relatively linear) correlation (although unsure how strong) with each other and the ‘O’ named lakes seem to have one as well with each other. The year variable does not appear to have a relationship to any of the variables. Also, it seems that the stream run-off variable has a relationship to the ‘O’ named lakes but no real notable relationship to the ‘A’ named lakes.

## Question 4

```{r}
# load dataset, select variables, preview dataset
data(Rateprof)

Rateprof <- Rateprof %>%
  select(c(quality, clarity, helpfulness, easiness, raterInterest))  

head(Rateprof)
```
```{r}
pairs(Rateprof)

```
Referring to the scatterplot matrix of the average professor ratings for the topics of quality, clarity, helpfulness, easiness, and rater interest, the variables quality, clarity, and helpfulness appear to each have strong positive correlations with each other. The variable easiness appears to have a much weaker positive correlation with helpfulness, clarity, and quality. 

##Question 5

# load and preview data
```{r}

data(student.survey)

head(student.survey)
```

##a
```{r}



ra <- student.survey %>%
  select(re, pi)

ggplot(data = ra) +
  geom_bar(mapping = aes(x = re, fill = pi)) +
  labs(title = "Relationship between Religiosity and Political Ideology", x = "Religiosity (how often you attend services)", y = "Political Ideology (pi)")
```
```{r}
tvhi <- student.survey %>%
  select(tv, hi)

ggplot(data = tvhi) +
  geom_point(mapping = aes(x = tv, y = hi)) +
  labs(title = "Relationship between Hours Watching TV and GPA", x = "Average Hours of TV watched per Week", y = "High School GPA")
```

##b

# inferential analysis: religiosity and political ideology
```{r}
lmrepa <- lm(data = student.survey, formula = as.numeric(pi) ~ as.numeric(re))
summary(lmrepa)
```
# looking at Pearson's correlation
```{r}
cor.test(as.numeric(student.survey$re), as.numeric(student.survey$pi))

```

# inferential analysis: hours of tv and high school gpa
```{r}
lmtvhi <- lm(data = student.survey, formula = hi ~ tv)
summary(lmtvhi)
```
# looking at Pearson's correlation
```{r}
cor.test(student.survey$tv, student.survey$hi)
```
 The people who attended religious services most weeks/every week were more likely to be conservative than those who never did.
 Also, there was no significant difference in political ideology between those who occasionally attended religious services and those who never did.
 
 Watching less hours of TV per week was associated with higher high-school GPA's.-That being said, as the R^2^ is fairly low, hours of TV watching is not a great predictor of high school GPA.


