---
title: "HW1"
author: "Niyati Sharma"
desription: "The first homework on descriptive statistics and probability"
date: "10/03/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - Niyati Sharma
  - desriptive statistics 
  - probability
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(readxl)
library(ggplot2)
library(stats)



knitr::opts_chunk$set(echo = TRUE)


```

## Question 1

## Read the data from xls file
```{r}
RE <- read_excel("_data/LungCapData.xls")
RE

```

##A

```{r}
RE %>% 
  ggplot(aes(LungCap))+
  geom_histogram(bins=20)
```

The histogram looks close to normal distributed.


## B


```{r}
RE %>%
  ggplot(aes (LungCap, color=Gender)) +
  geom_boxplot() +
  theme_classic() 
```
  The probability density of the female is higher than the males.


## C

```{r}
Mean_Smoker <- RE %>%
  group_by(Smoke) %>%
  summarise(mean = mean(LungCap))
Mean_Smoker

ggplot(RE, aes(LungCap,Smoke))+
  geom_boxplot()
```
From this sample, it appears that smokers have a higher mean lung capacity than non-smokers. 

## D

```{r}
RE<-RE %>% 
  mutate(Category = as.factor(case_when(Age <= 13 ~ "13 and under", 
                           Age == 14 |Age ==15 ~ "14-15", 
                           Age == 16 | Age==17 ~ "16-17",
                           Age >= 18 ~ "18 or over"
                           )))
RE

RE %>%
  ggplot(aes( LungCap, color = Smoke)) +
  geom_histogram()+
  facet_grid(Smoke ~ Category)
```
The people who smoke are few in age group of "less than or equal to 13".
From the result we can say  age is inversely proportional to the lung capacity.



## E
Form the above data we can say the output are pretty similar that smokers have a lower lung capacity compared to non-smokers

## F
correlation and covariance between lung capacity and age
```{r}
cov(RE$LungCap,RE$Age)
cor(RE$LungCap,RE$Age)

```
Covariance is positive and indicates that age and lung capacity are directly related.
Correlation is also positive,from these results we can conclude that the lung capacity increases with age. 


## Question 2

```{r}
x <- c(0:4)
freq <- c(128, 434, 160, 64, 24)
convictions <- data_frame(x, freq)
convictions

```

```{r}
convictions <- convictions %>% mutate(probability = freq/sum(freq))
convictions
```


## A
Probability of exactly 2 is 19.75%

## B
```{r}
a <-head(convictions,2)
sum(a$probability)
```
Probability that a randomly selected inmate has fewer than 2 prior
convictions : 69.38%

## C 

```{r}
a <-head(convictions,3)
sum(a$probability)
```
The probability that a randomly selected inmate has 2 or fewer prior
convictions : 89.13%

## D

```{r}
a <-tail(convictions,2)
sum(a$probability)
```
The probability that a randomly selected inmate has more than 2 prior
convictions? : 10.86%


## E
```{r}
WE <- weighted.mean(convictions$x,convictions$probability)
WE
```
The expected value for the number of prior convictions : 1.28

## F
The variance is 0.857 and the standard deviation is 0.925
```{r}

AB <- (sum(freq*((x-WE)^2)))/(sum(freq)-1)
AB
sqrt(AB)
```