---
title: "Final Project Proposal"
author: "Niyati Sharma"
description: Initial proposal for my final project
date: "10/11/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalpart1
  - Niyati Sharma
---

```{r}
library(tidyverse)
library(dplyr)

library(ggplot2)

knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

Credit risk is defined as the risk of loss resulting from the failure by a borrower to repay the principal and interest owed to the leader.So the purpose of credit analysis is to determine the creditworthiness of borrowers by measuring the risk of loss that the lender is exposed to.When calculating the credit risk of a particular borrower, lenders consider various factors like analyze different documents, such as the borrower’s income statement, balance sheet, credit reports, and other documents that reveal the financial situation of the borrower. to evaluate the characteristics of the borrower and conditions of the loan to estimate the probability of default and the subsequent risk of financial loss.

## Research Question

Q1. How credit risk depends on the age of the person.
Q2. Dominating factor on which credit risk depends.
Q3. Is credit risk depends on loan_intent?

## Hypothesis
According to research credit risk of a particular borrower, lenders consider various factors include the borrower’s capacity to repay are income, character, house ownership, and credit history.
Check the relationship between the age, income with credit risk with new dataset.


## Dataset
This dataset contains columns simulating credit bureau data, factors on which credit risk depends.
The variables of interest for me are income, age, employment length and home ownership.

```{r}
library(readr)
df <- read_csv("C:/Users/Lenovo/Downloads/credit_risk_dataset_1.csv")
summary(df)


```


