---
title: "Project Rough Draft Proposal"
author: "Yakub Rabiutheen"
description: "International Trade's influence on War"
date: "10/11/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - finalproject1
  - desriptive statistics 
  - probability
---

# Research Question

How has international trade influenced how countries interact with each other? This research project looks specifically at  France and Britain which  are grouped together as Colonial Powers to explore the relationship of Colonialism and international trade. This research project will be looking at data from the Correlates Of War Project, which has international trade data from 1870 to 2015. The cut-off year for this research project will be 1960, as  on December 14,1960, the UN declared Colonialism was a human's right's violation and legally declared Colonialism was over(McWhinney,1960).   


# Hypopthesis

This research project will be testing two hypothesis regarding Britain and France.

#Colonial Powers Hypopthesis.
1. The Years that France and Britain had more Exports is when the rate of colonization increased.
2. The Years that France and Britain had more Iron Production correlates to the years France and Britain increased levels of colonization.



```{r}
library(readxl)
library(readr)
Colonial_Years <- read_excel("C:/Users/yakub/Documents/GitHub/603_Fall_2022/posts/_data/Colonial_transformation_data.xls")
Imports_Exports<-read_csv("C:/Users/yakub/Documents/GitHub/603_Fall_2022/posts/_data/Countries_Imports_Exports.csv")
military_raw_metals<-read_csv("C:/Users/yakub/Documents/GitHub/603_Fall_2022/posts/_data/NMC_v4_0.csv")
```
#Descriptive Statistics.


As shown below, I tried finding Data regarding when colonialism began by France and Uk and seeing whether France and UK had more Trade Surpluses as they expanded their colonial empire. However, I was proven wrong as it appears that the UK has been running a Trade Deficit and has never had a Trade Surplus during their Colonial era pre-1960s. As such, I will have to change the approach of this research study. It appears the Balance of Trade has no relationship to Colonialism.









```{r}
colnames(Colonial_Years)[3] <- "Colonizing Country"
colnames(Colonial_Years)[4]<- "Year_Colonization_Began"
```


```{r}
library(dplyr)
Colonial_Years<-select(Colonial_Years,"Colonizing Country","Year_Colonization_Began")
Colonial_Years<-filter(Colonial_Years,`Colonizing Country` %in% c("F","UK"))
table(Colonial_Years)
```


```{r}
Imports_Exports%>% filter(year < '1960') 
```


```{r}
colonial_trade<-filter(Imports_Exports,`stateabb` %in% c("FRN","UKG"))

```










```{r}
options(scipen = 999)    
```


Created a Forumula to calculate Trade Surplus and Deficits.

```{r}
colonial_trade$trade_balance<-(colonial_trade$exports-colonial_trade$imports)
```




Found a better way to find years that France and Britain were running Trade Deficits.
```{r}
print(colonial_trade[colonial_trade$exports < colonial_trade$imports,] )
```


I did the inverse to find that the UK has always had a Trade Deficit
```{r}
print(colonial_trade[colonial_trade$exports > colonial_trade$imports,] )
```

My finding has found that there is no relationship between Trade Deficits and Colonialism as the UK has never had a positive trade balance.




##Conclusion

I think that the approach of my research has to be changed as my initial theory about trade deficits and Colonialism has been disapprove. As such, I think I will shift this project towards a different approach. I will try exploring the historical prices of commodity goods when France and U.K. were colonial powers.





# References
McWhinney, E. (1960, December 14). Declaration on the granting of Independence to colonial countries and Peoples. United Nations. Retrieved October 10, 2022, from https://legal.un.org/avl/ha/dicc/dicc.html 
```









































