---
title: "Homework 1 Solution"
author: "Dane Shelton"
description: "Intro to Quant Analysis Homework 1"
date: "10/1/22"
editor: visual
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

# Homework 1 Tasks:

1.) Using LungCapData, answer descriptive questions about the data and its distributions.

2.) Use the given distribution to answer questions about the probability of discrete events.

```{r}
#| label: Setup
#| include: false

#Loading Libraries
library(tidyverse)
library(summarytools)
```

## Task 1 - LungCapData

### Loading the Data

```{r}

#| include: false
#| label: Loading in LungCap

 og_lungcap <- readxl::read_xls("_data/LungCapData.xls")

# Quick look at dataset
# glimpse(og_lungcap)

# Variables - 3<dbl> ratio 3<char> (can coerce to logical if needed), 

# length(which(is.na(og_lungcap)))

# No missing values to consider

# Descriptive
# summarytools::dfSummary(og_lungcap)

```

LungCapData: Describes the lung capacity of a population of 725 children aged 3 - 19. It further categorizes the subjects by height, sex, smoking habits, and whether they were birthed using the Caesarean section technique.

In the following sections, we'll use `select()`, `group_by()`, `filter()`, and `summarize()` to further explore the data and find important relations between variables.

### a) LungCap Histogram

```{r}
#| label: Histogram of LungCap
#| echo: false
#| output: true

# Using GGPLOT to plot a histogram of variable - LungCap

ggplot(og_lungcap, aes(x=LungCap, y=..density..)) + 
  geom_histogram(bins=20) + 
  geom_vline(aes(xintercept=mean(LungCap)))
```

`LungCap` looks to be approximately normally distributed (unimodal, symmetric) with most observations centered around the mean (7.86).

### b) LungCap Histogram: by Gender

```{r}
#| label: Histogram by Gender
#| include: true

hist_gender <- ggplot(og_lungcap, aes(x=LungCap, y=..density.., fill=Gender)) +
  geom_histogram(alpha=.5, position="identity", bins=20)+
  geom_vline(aes(xintercept=mean(LungCap)))
hist_gender

```

Package ggplot2 functions `ggplot()` and `geom_histogram()` are used to display the `LungCap` distribution filled by the `Gender` variable. Both density plots center on the mean, indicating both male and female lung capacity observations are highly concentrated around the mean. The male distribution is shifted slightly to the right of the female distribution, meaning male observations had a higher upper range value than female observations. Males had more observations concentrated to the right of the mean, and the female distribution reciprocated this effect to the left of the mean.

### c) Smoking and Lung Capacity

```{r}
#| label: Smoker v Non Smoker group_by
#| output: true

smokers <- group_by(og_lungcap, Smoke)
smokers %>%
  summarize(mean(LungCap))

```

After creating a new dataset `smokers` by using `group_by()` on our original data, `smokers` is piped into a `summarize()` call. The results surprisingly show that the smoking group had a higher mean lung capacity than the nonsmoking group. This is likely due to a mean age difference within the groups.

### d) e) Smoking and Lung Capacity: Within Age Groupings

```{r}
#| label: Age Groupings
#| include: true
#| warning: false

# Creating Age Groups Using Case When

smokers_age <- smokers %>%
  mutate(AgeGroup = case_when(Age >= 18 ~ "18+", 
            Age == 16 | Age == 17 ~ "16-17",
            Age == 14 | Age == 15 ~ "14-15",
            Age <= 13~ "Under 13"))

# Mean LungCap by Age and Smoke
# Must regroup by Smoke again
smokers_age %>%
  group_by(AgeGroup, Smoke) %>%
    summarize(mean(LungCap))
```

After using `mutate()` to add a column `AgeGroup` to a copy of `smokers`, `group_by()` groups the new dataset by `AgeGroup` and `Smoke` before piping it into a `summarize()` command to find the grouped means of `LungCap` by `AgeGroup` *and* `Smoke`.

The results show that for children above the age of 13, smokers had a lower mean lung capacity than non-smokers. However, for the 13 and under group, we again see results that imply smokers have greater lung capacity than nonsmokers. Let's investigate further into the relationship between age and lung capacity to explain this quizzical result.

### f) Lung Capacity and Age

```{r}
#| label: Corr()and Cov() of Lung Capacity and Age
#| output: true

cov(og_lungcap$Age, og_lungcap$LungCap)
cor(og_lungcap$Age, og_lungcap$LungCap)

#GGPlot of Age vs Lung
ggplot(og_lungcap, aes(x=Age, y=LungCap)) + geom_point()
```

Age and LungCap have a high covariance which leads to a high correlation (p=0.82). This strong positive value (-1\<p\<1) indicates these variables "vary greatly" together: when Age is high in the data, so is LungCap. We cannot say that an increase in Age causes an increase Lung capacity without first showing this through regression; however, our results show the variables are highly correlated.

We can use knowledge of the human body to infer that as our body ages, our lungs mature. The ages of smokers of the Under 13 group are likely highly left skewed, as I don't expect many children under 10 to be smoking. This underlying age distribution explains our puzzling results from the previous section.

```{r}
#| label: Age Means withing Aged Smoking Groups
#| output: true
smokers_age%>%
  group_by(AgeGroup, Smoke) %>%
    summarize(mean(Age))
```

## Task 2 - Probability Distribution

First, let's create two vectors: x_val and freq. Then, we'll use rbind() to create a table.

```{r}
#| label: Creating Vectors and Binding
#| output: true

x_val <-c(0,1,2,3,4)
freq <- c(128,434,160,64,24)
prob <- freq/sum(freq)

xdist <- rbind(x_val,prob)

xdist

```

### a) P(x=2)

```{r}
#| label: a) P(x=2)
#| output: true
#| warning: false

# Finding probability of inmate having exactly 2 prior convictions

#Column Index is 3 as the first column is 0

#Surely there is a cleaner way to do this using tidyverse functions rather than base?

# a
a <- xdist['prob',3] 
a
```

### b) P(X\<2)

```{r}
#| label: b) P(x<2)
#| output: true
#| warning: false

#b
b <- sum(xdist['prob',1:2])
b
```

### c) P(x\<=2)

```{r}
#| label: c) P(x<=2)
#| output: true
#| warning: false

# c
c <- a + b
c
```

### d) P(x\>2)

```{r}
#| label: d) P(x>2)
#| output: true
#| warning: false

#d
d <- 1 - c
d
```

### e) E(X)

```{r}
#| label: E(X)
#| output: true
#| echo: false

ex <- sum(x_val*prob)
ex

```

### f) Variance and SD

#### Variance

```{r}
#| label: Var
#| output: true
#| echo: true

# Var= E(X^2) - E(X)^2
# Again using brute force because cannot use var() function on the object xdist correctly
var_x <-sum((x_val^2)*prob) - ex^2
var_x
```

#### Standard Deviation

```{r}
#| label: std
#| output: true
#| echo: false

# SD = sqrt(var)
sd <- sqrt(var_x)
sd
```
