---
title: "ShoshanaBuck-HW1"
author: "Shoshana Buck"
description: "First homeowrk on descriptive statistics and probability"
date: "10/03/22"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - desriptive statistics 
  - probability
---

```{r}
#install libraries
library(tidyverse)
library(readxl)
library(ggplot2)
```

# Question 1

Use the LungCapData to answer the following questions.

```{r}
# read in data
lung_cap<- read_xls("_data/LungCapData.xls")
lung_cap
```

### a. What does the distribution of LungCap look like?

```{r}
#plot histogram
hist(lung_cap$LungCap)
```

The histogram shows that the distribution is pretty close to a normal distribution, with an almost a bell shaped curve. Meaning that the data near the mean are more of a frequent occurrence which is true because there are fewer observations near the margins.

### b. compare the probability distribution of the LungCap with respect to Males and Females?

```{r}
#Box plot
ggplot(lung_cap, aes (Gender,LungCap)) + geom_boxplot()
```

The box plot shows that male's have a slightly higher lung capacity than females. Female's have more values in the first quartile range and a lower minimum value than male's. On the other hand male's have a higher max value and more values in the 3rd quartile range.

### c. Compare the mean lung capacities for smokers and non-smokers. Does it make sense?

```{r}
lung_cap %>% 
  group_by(Smoke) %>% 
  summarise(avg_lung_cap = mean(LungCap))


```

This does not make sense. Smokers have a higher sample mean than non-smokers which intuitively does not make sense because we would assume non-smokers would have a higher lung capacity.

### d. Examine the relationship between Smoking and Lung Capacity within age groups: "less than or equal to 13", "14 to 15", "16 to 17", and "greater than or equal to 18".

```{r}
#categorical variable of age_groups 
df<- lung_cap %>% 
group_by(Smoke,LungCap) %>% 
  summarise(age_group = case_when(Age<=13 ~ "13 & under",Age ==14 | Age == 15 ~ "14 to 15",Age== 16 | Age == 17 ~ "16 to 17", Age>=18 ~ "18 & older")) 

#mean of lung capacity with new variable
df %>% 
  group_by(Smoke, age_group) %>% 
  summarise(avg_lung_cap = mean(LungCap)) %>% 
  arrange(desc(avg_lung_cap))
#histogram
ggplot(df, aes(x = LungCap)) + facet_grid(Smoke ~age_group) +geom_histogram()
```

Using the package ***ggplot*** I used the function *facet_grids* to show a good visualization of the lung capacity between non-smokers and smokers within each age group. Looking at the histograms all *age_groups* that are non-smokers have a higher sample mean proving that non-smokers have

### e. Compare the lung capacities for smokers and non-smokers within each age group. Is your answer different from the one in part d. What could possibly be going on here?

```{r}
# Mean of non-smokers 13 & younger
df %>% 
  filter(Smoke == 'no' & age_group == '13 & under') %>% 
  pull(LungCap) %>% 
  mean()
# Mean of smokers 13 & younger
df %>% 
  filter(Smoke == 'yes' & age_group == '13 &under') %>% 
  pull(LungCap) %>% 
  mean()
#Mean of non-smokers 14 to 15
df %>% 
  filter(Smoke == 'no' & age_group == '14 to 15') %>% 
  pull(LungCap) %>% 
  mean()
#Mean of smokers 14 to 15
df %>% 
  filter(Smoke == 'yes' & age_group == '14 to 15') %>% 
  pull(LungCap) %>% 
  mean()
#Mean of non-smokers 16 to 17
df %>% 
  filter(Smoke == 'no' & age_group == '16 to 17') %>% 
  pull(LungCap) %>% 
  mean()
#Mean of smokers 16 to 17
df %>% 
  filter(Smoke == 'yes' & age_group == '16 to 17') %>% 
  pull(LungCap) %>% 
  mean()
#Mean of non-smokers 18 & older
df %>% 
  filter(Smoke == 'no' & age_group == '18 & older') %>% 
  pull(LungCap) %>% 
  mean()
# Mean of smokers 18 & older
df %>% 
  filter(Smoke == 'yes' & age_group == '18 & older') %>% 
  pull(LungCap) %>% 
  mean()
lung_cap

```

### f. Calculate the correlation and covariance between Lung Capacity and Age. (use the cov() and cor() functions in R). Interpret your results.

```{r}
#Correlation
cor(lung_cap$LungCap, lung_cap$Age)
#Covariance
cov(lung_cap$LungCap, lung_cap$Age)
```

The correlation is 0.81 which is pretty close to +1, meaning that the there is a positive relationship between lung capacity and age. The covariance is also high which shows that the two variables of lung capacity and age have a positive relationship. #2 Let X = number of prior convictions for prisoners at a state prison at which there are 810 prisoners.

```{r}
# x= prior convictions 
x<-c(0, 1, 2, 3, 4)
frequency<-c(128, 434, 160, 64, 24)
state_prison<- data_frame(x,frequency)
state_prison
```

### a. What is the probability that a randomly selected inmate has exactly 2 prior convictions?

```{r}
# 2 prior convictions.P(2)/total
160/810

```

There is a 1.9% probability that a randomly selected inmate has exactly 2 prior convictions.

### b. What is the probability that a randomly selected inmate has fewer than 2 prior convictions?

```{r}
#  less than 2 prior convictions. (P(0) + P(1))/total 
(128+434)/810

```

There is a 6.9% probability that a randomly selected inmate has fewer than 2 prior convictions.

### c. What is the probability that a randomly selected inmate has 2 or fewer prior convictions?

```{r}
# 2 or fewer prior convictions. (P(0) + P(1) + P(2)) +total
(128+434+160)/810
```

There is a 8.9% probability that a randomly selected inmate has 2 or fewer prior convictions.

### d. What is the probability that a randomly selected inmate has more than 2 prior convictions?

```{r}
# More than 2 prior convictions. (P(3) +P(4)) + total
(64+24)/810

```

There is a 10.8% probability that a randomly selected inmate has more than 2 prior convictions.

### e. What is the expected value for the number of prior convictions?

```{r}
#Prior convictions. ((0*P(0)) +(1*(P(1)) + (2*P(2)) + (3*P(3)) + (4*P(4)))
df<-((128*0/810) +(434*1/810) +(160*2/810) +(64*3/810) +(24*4/810)) 
mean(df)
```

The expected value for number of prior convictions is 1.28

### f. Calculate the variance and the standard deviation for the Prior Convictions.

```{r}
# variance
v<- ((0-1.28)^2) *(128/810) +((1-1.28)^2) * (434/810)+((2-1.28)^2) * (160/810)+((3-1.28)^2) *(64/810) +((4-1.28)^2) * (24/810)
v
# standard deviation
sd<- sqrt(v)
sd
```

The variance is 0.856 and the standard deviation is 0.925.
