---
<<<<<<< HEAD
title: "Graphs and Probz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(ggplot2)
library(markdown)
library(rmarkdown)
library(tidyr)
library(tidyselect)
library(tidyverse)
```


```{r}
library(readxl)
LungCapData <- read_excel("_data/LungCapData.xls")
View(LungCapData)                                                              
```


```{r} 
m_lung<-mean(LungCapData$LungCap)
sd_lung<-sd(LungCapData$LungCap)

hist(LungCapData$LungCap, prob= TRUE, xlim = c(0, 20))
curve(dnorm(x, m_lung, sd_lung), add= TRUE,lwd= 2,col= "blue")

```
# Section 1 Question 1
Looks like a normal distribution


### Question 2

```{r}
grouped_gender<- LungCapData %>% group_by(Gender)
summarize(grouped_gender)
qplot(data = grouped_gender, x = Gender, y = LungCap, geom = "boxplot")
```
Males have a higher mean than females. 

### Question 3
Smokers have a higher mean
```{r}
grouped_smokers<- LungCapData %>% group_by(Smoke)
summarize(grouped_smokers)
qplot(data = grouped_smokers, x = Smoke, y = LungCap, geom = "boxplot")
```


### Question 4
Looks like the lung capacity is highest for children ages 0-13, specifically for males. 
```{r}
LungCapData$Agegroups<-cut(LungCapData$Age,breaks=c(-Inf, 13, 15, 17, 20), labels=c("0-13 years", "14-15 years", "16-17 years", "18+ years"))

ggplot(LungCapData, aes(x = LungCap, y = Agegroups, fill = Gender)) +
          geom_bar(stat = "identity") +
          coord_flip() +
          theme_classic()
                
```

## Question 5
Doesnt look like its good being a smoker under the age of 18, or any age. Lung capacity is smaller for these groups
```{r}
ggplot(LungCapData, aes(x = LungCap, y = Agegroups, fill = Smoke)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    theme_classic()

```


## Question 6

```{r}
covar<-cov(LungCapData$LungCap, LungCapData$Age)
print(covar)
corre<-cor(LungCapData$LungCap, LungCapData$Age, method = "pearson")
print(corre)
```


# Section 2 Question 2
 
```{r}
a<-128/810
a
b<-434/810
b
c<-160/810
c
d<-64/810
d
e<-24/810
e
ei<-((a*0)+(b*1)+(c*2)+(d*3)+(e*4))
ei
varei<-((0-ei)^2+(1-ei)^2+(2-ei)^2+(3-ei)^2+(4-ei)^2)/5
varei
sdei<-sqrt(varei)
sdei
```




=======
title: "Blog Post Template"
author: "Your Name"
desription: "Something to describe what I did"
date: "08/02/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw1
  - challenge1
  - my name
  - dataset
  - ggplot2
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

This document provides yaml header inforamtion you will need to replicate each week to submit your homework or other blog posts. Please observe the following conventions:

-   Save your own copy of this template as a blog post in the `posts` folder, naming it `FirstLast_hwX.qmd`
-   Edit the yaml header to change your author name - use the same name each week
-   include a *description* that is reader friendly
-   update the *category* list to indicate the type of submission, the data used, the main packages or techniques, your name, or any thing else to make your document easy to find
-   edit as a normal qmd/rmd file


```{r}
x <- c(2,3,4,5)
mean(x)
```

## Rendering your post

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. 

:::{.callout-warning}
Be sure that you have moved your `*.qmd` file into the `posts` folder **BEFORE** you render it, so that all files are stored in the correct location.
:::

:::{.callout-important}
Only render a single file - don't try to render the whole website!
:::

:::{.callout-note}
## Pilot Student Blogs

We are piloting a workflow including individual student websites with direted and limited pull requests back to course blogs. Please let us know if you would like to participate.
:::

## Reading in data files

The easiest data source to use - at least initially - is to choose something easily accessible, either from our `_data` folder provided, or from an online source that is publicly available.

:::{.callout-tip}
## Using Other Data
If you would like to use a source that you have access to and it is small enough and you don't mind making it public, you can copy it into the `_data` file and include in your *commit* and *pull request*.
:::

:::{.callout-tip}
## Using Private Data

If you would like to use a proprietary source of data, that should be possible using the same process outlined above. There may initially be a few issues. We hope to have this feature working smoothly soon!
>>>>>>> 3c1aae64bdd557bd046b3c1aa77e8ed147b948a5
